---
title: "alkati_subsampling_simulations_2"
author: "Haider Inam"
date: "10/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=rprojroot::find_rstudio_root_file())
```


```{r,include=F}
# rm(list=ls())
#Loading required packages
library(knitr)
library(tictoc)
library(workflowr)
library(VennDiagram)
library(dplyr)
library(foreach)
library(doParallel)
library(ggplot2)
library(reshape2)
library(RColorBrewer)
library(devtools)
library(ggsignif)
library(plotly)
library(BiocManager)
######################Cleanup for GGPlot2#########################################
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        # axis.line = element_line(color = "black"),
        axis.text = element_text(face="bold",color="black",size="11"),
        text=element_text(size=11,face="bold"),
        axis.title=element_text(face="bold",size="11"))
```

```{r}

source("code/contab_maker.R")
source("code/contab_simulator.R")
source("code/contab_downsampler.R")
source("code/alldata_compiler.R")
source("code/mut_excl_genes_generator.R")

# source("../code/contab_maker.R")
# source("../code/contab_simulator.R")
# source("../code/contab_downsampler.R")
# source("../code/alldata_compiler.R")
# source("../code/mut_excl_genes_generator.R")


```
  
  
### Section 1: Pairwise Comparison of ALKATI BRAF vs BRAF NRAS (next section compares ALKATI NRAS vs BRAF NRAS)
```{r}
  nameposctrl1<-'BRAF'
  #Positive control 1
  nameposctrl2<-'NRAS'
  #Oncogene in Question
  namegene<-'ATI'
  #Mutation Boolean (Y or N)
  mtn<-'N'
  #Name Mutation for Positive Ctrl 1
  nameposctrl1mt<-'V600E'
  #Name of Mutation for Positive Ctrl 2
  nameposctrl2mt<-'Q61L'

alldata=read.csv("output/all_data_skcm.csv",sep=",",header=T,stringsAsFactors=F)
# alldata=read.csv("../output/all_data_skcm.csv",sep=",",header=T,stringsAsFactors=F)
head(alldata)
```


```{r}
# rm(list=ls())
###Not mutation specific generation of counts###
alldata_comp=alldata_compiler(alldata,nameposctrl1,nameposctrl2,namegene,'N',"N/A","N/A")[[2]]
head(alldata_comp)

###Calculating Odds ratios and GOI frequencies for the raw data###
cohort_size=length(alldata_comp$Positive_Ctrl1)
pc1pc2_contab_counts=contab_maker(alldata_comp$Positive_Ctrl1,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
# pc1pc2_contab_counts=pc1new_pc2_contab

goipc1_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl1,alldata_comp)[2:1, 2:1]
# goipc1_contab_counts=goinew_pc1_contab
###Had to add the 2:1 bits because the contab maker spits out NN YY whereas we wanted YNYN
goipc2_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
# pc1pc2_contab_counts=gene_pair_2_table
# goipc1_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl1,alldata_comp)[2:1, 2:1]
# goipc1_contab_counts=gene_pair_1_table

cohort_size_curr=cohort_size
# goipc2_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
pc1pc2_contab_probabilities=pc1pc2_contab_counts/cohort_size_curr
goipc1_contab_probabilities=goipc1_contab_counts/cohort_size_curr
goipc2_contab_probabilities=goipc2_contab_counts/cohort_size_curr
# pc1pc2_contab_probabilities=pc1pc2_contab_counts
# goipc1_contab_probabilities=goipc1_contab_counts
# goipc2_contab_probabilities=goipc2_contab_counts/cohort_size

or_pc1pc2=pc1pc2_contab_probabilities[1,1]*pc1pc2_contab_probabilities[2,2]/(pc1pc2_contab_probabilities[1,2]*pc1pc2_contab_probabilities[2,1])
or_goipc1=goipc1_contab_probabilities[1,1]*goipc1_contab_probabilities[2,2]/(goipc1_contab_probabilities[1,2]*goipc1_contab_probabilities[2,1])
or_goipc2=goipc2_contab_probabilities[1,1]*goipc2_contab_probabilities[2,2]/(goipc2_contab_probabilities[1,2]*goipc2_contab_probabilities[2,1])

goi_freq=goipc1_contab_probabilities[1,1]+goipc1_contab_probabilities[1,2]
# goi_freq=.25
# class(goi_freq)

###

###Downsampling PC1 to the probability of GOI without changing ORs###
###The function below converts contingency table data to a new contingency table in which the data is downsampled to the desired frequency, aka the frequency of the GOI in this case###
pc1new_pc2_contab=contab_downsampler(pc1pc2_contab_probabilities,goi_freq)
goinew_pc1_contab=contab_downsampler(goipc1_contab_probabilities,goi_freq)
goinew_pc2_contab=contab_downsampler(goipc2_contab_probabilities,goi_freq)
##original contab:
head(pc1pc2_contab_probabilities)
###downsampled contab:
head(pc1new_pc2_contab)

pc1rawpc2_contabs_sims=contab_simulator(pc1pc2_contab_probabilities,1000,cohort_size_curr)
pc1pc2_contabs_sims=contab_simulator(pc1new_pc2_contab,1000,cohort_size_curr)
goipc1_contabs_sims=contab_simulator(goinew_pc1_contab,1000,cohort_size_curr)
goipc2_contabs_sims=contab_simulator(goinew_pc2_contab,1000,cohort_size_curr)
# goipc2_contabs_sims=contab_simulator(goinew_pc2_contab,1000,cohort_size)
# head(pc1pc2_contabs_sims) #each row in this dataset is a new contab
pc1rawpc2_contabs_sims=data.frame(pc1rawpc2_contabs_sims)
  pc1rawpc2_contabs_sims=pc1rawpc2_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))

pc1pc2_contabs_sims=data.frame(pc1pc2_contabs_sims)
  pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))
  
goipc1_contabs_sims=data.frame(goipc1_contabs_sims)
  goipc1_contabs_sims=goipc1_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))
  
goipc2_contabs_sims=data.frame(goipc2_contabs_sims)
goipc2_contabs_sims=goipc2_contabs_sims%>%
mutate(or=p11*p00/(p10*p01))
  
# goipc2_contabs_sims=data.frame(goipc2_contabs_sims)
#   goipc2_contabs_sims=goipc2_contabs_sims%>%
#   mutate(or=p11*p00/(p10*p01))
pc1rawpc2_contabs_sims$comparison="pc1rawpc2"
pc1pc2_contabs_sims$comparison="pc1pc2"
goipc1_contabs_sims$comparison="goipc1"
goipc2_contabs_sims$comparison="goipc2"
or_median_raw=quantile(pc1rawpc2_contabs_sims$or,na.rm = T)[3]
or_uq_raw=quantile(pc1rawpc2_contabs_sims$or,na.rm = T)[4]
or_median_downsampled=quantile(pc1pc2_contabs_sims$or,na.rm = T)[3]
or_uq_downsampled=quantile(pc1pc2_contabs_sims$or,na.rm = T)[4]

pc1rawpc2_contabs_sims=pc1rawpc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
goipc1_contabs_sims=goipc1_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )

goipc2_contabs_sims=goipc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
# pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
# goipc1_contabs_sims=goipc1_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
# goipc2_contabs_sims=goipc2_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
pc1rawpc2_isgreater_raw_median=sum(pc1rawpc2_contabs_sims$isgreater_raw_median)
pc1rawpc2_isgreater_raw_uq=sum(pc1rawpc2_contabs_sims$isgreater_raw_uq)
pc1rawpc2_isgreater_median=sum(pc1rawpc2_contabs_sims$isgreater_median)
pc1rawpc2_isgreater_uq=sum(pc1rawpc2_contabs_sims$isgreater_uq)

pc1pc2_isgreater_raw_median=sum(pc1pc2_contabs_sims$isgreater_raw_median)
pc1pc2_isgreater_raw_uq=sum(pc1pc2_contabs_sims$isgreater_raw_uq)
pc1pc2_isgreater_median=sum(pc1pc2_contabs_sims$isgreater_median)
pc1pc2_isgreater_uq=sum(pc1pc2_contabs_sims$isgreater_uq)

goipc1_isgreater_raw_median=sum(goipc1_contabs_sims$isgreater_raw_median)
goipc1_isgreater_raw_uq=sum(goipc1_contabs_sims$isgreater_raw_uq)
goipc1_isgreater_median=sum(goipc1_contabs_sims$isgreater_median)
goipc1_isgreater_uq=sum(goipc1_contabs_sims$isgreater_uq)


plotting_df=rbind(pc1pc2_contabs_sims,goipc1_contabs_sims,goipc2_contabs_sims)
# plotting_df=rbind(pc1pc2_contabs_sims,goipc1_contabs_sims)
# 


ggplot(plotting_df,aes(x=(or),fill=comparison))+
  geom_histogram(bins=40,alpha=0.55,position="identity")+
  # geom_histogram(bins=50,alpha=0.55)+
  scale_y_continuous(expand=c(0,0),name="Count")+
  scale_x_continuous(expand=c(0,0),trans="log10",name="Odds Ratio")+
  scale_fill_brewer(palette="Set2")+
  # geom_vline(xintercept = or_pc1pc2)+
  cleanup

ggplot(plotting_df,aes(y=(or),x=comparison),fill=factor(comparison))+
  geom_boxplot()+
  scale_y_continuous(name="Odds Ratio",trans="log10")+
  scale_x_discrete(name="")+
  scale_fill_brewer(palette="Set2")+
  geom_hline(yintercept = or_uq_downsampled,linetype="dashed")+
  cleanup+
  theme(legend.position = "none",
        axis.ticks.x = element_blank())

# ggsave("paircon_boxplot.pdf",width = 3,height=2,units="in",useDingbats=F)
```
  
    
### Section 1 Conclusion: ALKATI vs BRAF has no overlap with BRAF vs NRAS, i.e. 100% of the BRAF NRAS simulations lie outside of the ALKATI BRAF regime.


### Section 2: Pairwise Comparison of ALKATI NRAS vs BRAF NRAS (the previous section compared ALKATI BRAS vs BRAF NRAS)
```{r}
  nameposctrl1<-'NRAS'
  #Positive control 1
  nameposctrl2<-'BRAF'
  #Oncogene in Question
  namegene<-'ATI'
  #Mutation Boolean (Y or N)
  mtn<-'N'
  #Name Mutation for Positive Ctrl 1
  nameposctrl1mt<-'Q61L'
  #Name of Mutation for Positive Ctrl 2
  nameposctrl2mt<-'V600E'

alldata=read.csv("output/all_data_skcm.csv",sep=",",header=T,stringsAsFactors=F)
# alldata=read.csv("../output/all_data_skcm.csv",sep=",",header=T,stringsAsFactors=F)
head(alldata)
```



```{r}
# rm(list=ls())
###Not mutation specific generation of counts###
alldata_comp=alldata_compiler(alldata,nameposctrl1,nameposctrl2,namegene,'N',"N/A","N/A")[[2]]
head(alldata_comp)

###Calculating Odds ratios and GOI frequencies for the raw data###
cohort_size=length(alldata_comp$Positive_Ctrl1)
pc1pc2_contab_counts=contab_maker(alldata_comp$Positive_Ctrl1,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
# pc1pc2_contab_counts=pc1new_pc2_contab

goipc1_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl1,alldata_comp)[2:1, 2:1]
# goipc1_contab_counts=goinew_pc1_contab
###Had to add the 2:1 bits because the contab maker spits out NN YY whereas we wanted YNYN
goipc2_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
# pc1pc2_contab_counts=gene_pair_2_table
# goipc1_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl1,alldata_comp)[2:1, 2:1]
# goipc1_contab_counts=gene_pair_1_table

cohort_size_curr=cohort_size
# goipc2_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
pc1pc2_contab_probabilities=pc1pc2_contab_counts/cohort_size_curr
goipc1_contab_probabilities=goipc1_contab_counts/cohort_size_curr
goipc2_contab_probabilities=goipc2_contab_counts/cohort_size_curr
# pc1pc2_contab_probabilities=pc1pc2_contab_counts
# goipc1_contab_probabilities=goipc1_contab_counts
# goipc2_contab_probabilities=goipc2_contab_counts/cohort_size

or_pc1pc2=pc1pc2_contab_probabilities[1,1]*pc1pc2_contab_probabilities[2,2]/(pc1pc2_contab_probabilities[1,2]*pc1pc2_contab_probabilities[2,1])
or_goipc1=goipc1_contab_probabilities[1,1]*goipc1_contab_probabilities[2,2]/(goipc1_contab_probabilities[1,2]*goipc1_contab_probabilities[2,1])
or_goipc2=goipc2_contab_probabilities[1,1]*goipc2_contab_probabilities[2,2]/(goipc2_contab_probabilities[1,2]*goipc2_contab_probabilities[2,1])

goi_freq=goipc1_contab_probabilities[1,1]+goipc1_contab_probabilities[1,2]
# goi_freq=.25
# class(goi_freq)

###

###Downsampling PC1 to the probability of GOI without changing ORs###
###The function below converts contingency table data to a new contingency table in which the data is downsampled to the desired frequency, aka the frequency of the GOI in this case###
pc1new_pc2_contab=contab_downsampler(pc1pc2_contab_probabilities,goi_freq)
goinew_pc1_contab=contab_downsampler(goipc1_contab_probabilities,goi_freq)
goinew_pc2_contab=contab_downsampler(goipc2_contab_probabilities,goi_freq)
##original contab:
head(pc1pc2_contab_probabilities)
###downsampled contab:
head(pc1new_pc2_contab)

pc1rawpc2_contabs_sims=contab_simulator(pc1pc2_contab_probabilities,1000,cohort_size_curr)
pc1pc2_contabs_sims=contab_simulator(pc1new_pc2_contab,1000,cohort_size_curr)
goipc1_contabs_sims=contab_simulator(goinew_pc1_contab,1000,cohort_size_curr)
goipc2_contabs_sims=contab_simulator(goinew_pc2_contab,1000,cohort_size_curr)
# goipc2_contabs_sims=contab_simulator(goinew_pc2_contab,1000,cohort_size)
# head(pc1pc2_contabs_sims) #each row in this dataset is a new contab
pc1rawpc2_contabs_sims=data.frame(pc1rawpc2_contabs_sims)
  pc1rawpc2_contabs_sims=pc1rawpc2_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))

pc1pc2_contabs_sims=data.frame(pc1pc2_contabs_sims)
  pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))
  
goipc1_contabs_sims=data.frame(goipc1_contabs_sims)
  goipc1_contabs_sims=goipc1_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))
  
goipc2_contabs_sims=data.frame(goipc2_contabs_sims)
goipc2_contabs_sims=goipc2_contabs_sims%>%
mutate(or=p11*p00/(p10*p01))
  
# goipc2_contabs_sims=data.frame(goipc2_contabs_sims)
#   goipc2_contabs_sims=goipc2_contabs_sims%>%
#   mutate(or=p11*p00/(p10*p01))
pc1rawpc2_contabs_sims$comparison="pc1rawpc2"
pc1pc2_contabs_sims$comparison="pc1pc2"
goipc1_contabs_sims$comparison="goipc1"
goipc2_contabs_sims$comparison="goipc2"
or_median_raw=quantile(pc1rawpc2_contabs_sims$or,na.rm = T)[3]
or_uq_raw=quantile(pc1rawpc2_contabs_sims$or,na.rm = T)[4]
or_median_downsampled=quantile(pc1pc2_contabs_sims$or,na.rm = T)[3]
or_uq_downsampled=quantile(pc1pc2_contabs_sims$or,na.rm = T)[4]

pc1rawpc2_contabs_sims=pc1rawpc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
goipc1_contabs_sims=goipc1_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )

goipc2_contabs_sims=goipc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
# pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
# goipc1_contabs_sims=goipc1_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
# goipc2_contabs_sims=goipc2_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
pc1rawpc2_isgreater_raw_median=sum(pc1rawpc2_contabs_sims$isgreater_raw_median)
pc1rawpc2_isgreater_raw_uq=sum(pc1rawpc2_contabs_sims$isgreater_raw_uq)
pc1rawpc2_isgreater_median=sum(pc1rawpc2_contabs_sims$isgreater_median)
pc1rawpc2_isgreater_uq=sum(pc1rawpc2_contabs_sims$isgreater_uq)

pc1pc2_isgreater_raw_median=sum(pc1pc2_contabs_sims$isgreater_raw_median)
pc1pc2_isgreater_raw_uq=sum(pc1pc2_contabs_sims$isgreater_raw_uq)
pc1pc2_isgreater_median=sum(pc1pc2_contabs_sims$isgreater_median)
pc1pc2_isgreater_uq=sum(pc1pc2_contabs_sims$isgreater_uq)

goipc1_isgreater_raw_median=sum(goipc1_contabs_sims$isgreater_raw_median)
goipc1_isgreater_raw_uq=sum(goipc1_contabs_sims$isgreater_raw_uq)
goipc1_isgreater_median=sum(goipc1_contabs_sims$isgreater_median)
goipc1_isgreater_uq=sum(goipc1_contabs_sims$isgreater_uq)


plotting_df=rbind(pc1pc2_contabs_sims,goipc1_contabs_sims,goipc2_contabs_sims)
# plotting_df=rbind(pc1pc2_contabs_sims,goipc1_contabs_sims)
# 


ggplot(plotting_df,aes(x=(or),fill=comparison))+
  geom_histogram(bins=40,alpha=0.55,position="identity")+
  # geom_histogram(bins=50,alpha=0.55)+
  scale_y_continuous(expand=c(0,0),name="Count")+
  scale_x_continuous(expand=c(0,0),trans="log10",name="Odds Ratio")+
  scale_fill_brewer(palette="Set2")+
  # geom_vline(xintercept = or_pc1pc2)+
  cleanup

ggplot(plotting_df,aes(y=(or),x=comparison),fill=factor(comparison))+
  geom_boxplot()+
  scale_y_continuous(name="Odds Ratio",trans="log10")+
  scale_x_discrete(name="")+
  scale_fill_brewer(palette="Set2")+
  geom_hline(yintercept = or_uq_downsampled,linetype="dashed")+
  cleanup+
  theme(legend.position = "none",
        axis.ticks.x = element_blank())

# ggsave("paircon_boxplot.pdf",width = 3,height=2,units="in",useDingbats=F)
```
  
    
### Section 2 Conclusion: ALKATI vs NRAS has no overlap with BRAF vs NRAS, i.e. 100% of the BRAF NRAS simulations lie outside of the ALKATI BRAF regime.
