  ---
title: "ALKATI Filter Cutoff Analysis"
author: "Haider Inam"
date: "2/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..") #To change the root directory from Analysis to medsproject
```


```{r}
library(ggplot2)
library(knitr)
library(dplyr)
library(tictoc)
library(foreach)
library(doParallel)
source("code/alldata_compiler.R")
source("code/contab_maker.R")
######################Cleanup for GGPlot2#########################################
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))
```

####**Varying Filter for mean RPKM from 1:100**
####**Note that by mean RPKM, I mean ratio of 20-29 RPKM/1-19 RPKM**
```{r}
#Run Filter Analysis
ct=1
simresults<-matrix(nrow=100,ncol=2)
alldata=read.csv("data/all_data.csv",sep=",",header=T,stringsAsFactors=F)
for (meanRPKM in 1:100){
  alldata_filtered=alldata%>%
    group_by(Patid,mean_RPKM_1.19,mean_RPKM_20.29,Ratio20.29, mRNA_count,BRAF,NRAS,RSEM_normalized)%>%
    summarize(ATI=as.numeric(mRNA_count>=500&Ratio20.29>meanRPKM&RSEM_normalized>=100)[1])
  
  alldata_comp=alldata_compiler(alldata_filtered,'BRAF','NRAS','ATI','N',"N/A","N/A")[[2]]

  contab_pc1_genex=contab_maker(alldata_comp$Positive_Ctrl1,alldata_comp$genex,alldata_comp)
  simresults[ct,1]=ct #Total Count
  simresults[ct,2]=fisher.test(contab_pc1_genex)$p.value #p.value
  ct=ct+1
}
count=simresults[c(1:100),1]
pval=simresults[c(1:100),2]
colnames(simresults)=c("totCt","p_val")
cols=c("totCt","p_val")

simresults=as.data.frame(simresults, stringsAsFactors = F, )
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.character)
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.numeric)

#Plot Results
ggplot(simresults,aes(x=totCt,y=p_val))+geom_line(size=3)+ggtitle("Varying RPKM Filter")+xlab("Exon20:29/Exon1:19")+ylab("P-value")+theme_bw()+theme(plot.title = element_text(hjust=.5),text = element_text(size=30,face="bold"),axis.title = element_text(face="bold",size="26"),axis.text=element_text(face="bold",size="26"))
  
# ggsave("filteranalysis_RPKM.pdf",width = 9,height = 6,units = "in",useDingbats=F)
```
####**Varying Filter for mean RSEM from 1 to 1000**
####**Note that these are the sum of RSEM across all ALK-exons**
```{r}
#Run Filter Analysis
ct=1
simresults<-matrix(nrow=1000,ncol=2)
alldata=read.csv("data/all_data.csv",sep=",",header=T,stringsAsFactors=F)
for (rsem in 1:1000){
  alldata_filtered=alldata%>%
    group_by(Patid,mean_RPKM_1.19,mean_RPKM_20.29,Ratio20.29, mRNA_count,BRAF,NRAS,RSEM_normalized)%>%
    summarize(ATI=as.numeric(mRNA_count>=500&Ratio20.29>10&RSEM_normalized>=rsem)[1])   
  alldata_comp=alldata_compiler(alldata_filtered,'BRAF','NRAS','ATI','N',"N/A","N/A")[[2]]
  
  contab_pc1_genex=contab_maker(alldata_comp$Positive_Ctrl1,alldata_comp$genex,alldata_comp)
  simresults[ct,1]=ct #Total Count
  simresults[ct,2]=fisher.test(contab_pc1_genex)$p.value #p.value
  ct=ct+1
}
count=simresults[c(1:100),1]
pval=simresults[c(1:100),2]
colnames(simresults)=c("totCt","p_val")
cols=c("totCt","p_val")

simresults=as.data.frame(simresults, stringsAsFactors = F, )
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.character)
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.numeric)

#Plot Results
ggplot(simresults,aes(x=totCt,y=p_val))+geom_line(size=3)+ggtitle("Varying RSEM Filter")+xlab("RSEM")+ylab("P-Value")+theme_bw()+theme(plot.title = element_text(hjust=.5),text = element_text(size=30,face="bold"),axis.title = element_text(face="bold",size="26"),axis.text=element_text(face="bold",size="26"))
  
# ggsave("filteranalysis_RSEM.pdf",width = 9,height = 6,units = "in",useDingbats=F)
```
####**Varying Filter for ALK raw read count from 1 to 10,000**
####**Note that these are the sum of mRNA counts across all ALK-exons**
```{r}
#Run Filter Analysis
ct=1
simresults<-matrix(nrow=1000,ncol=2)
alldata=read.csv("data/all_data.csv",sep=",",header=T,stringsAsFactors=F)
for (rawreadcount in 1:1000){
  alldata_filtered=alldata%>%
    group_by(Patid,mean_RPKM_1.19,mean_RPKM_20.29,Ratio20.29, mRNA_count,BRAF,NRAS,RSEM_normalized)%>%
    summarize(ATI=as.numeric(mRNA_count>=rawreadcount&Ratio20.29>10&RSEM_normalized>=100)[1])
  alldata_comp=alldata_compiler(alldata_filtered,'BRAF','NRAS','ATI','N',"N/A","N/A")[[2]]
  
  contab_pc1_genex=contab_maker(alldata_comp$Positive_Ctrl1,alldata_comp$genex,alldata_comp)
  
  simresults[ct,1]=ct #Total Count
  simresults[ct,2]=fisher.test(contab_pc1_genex)$p.value #p.value
  ct=ct+1
  }
count=simresults[c(1:100),1]
pval=simresults[c(1:100),2]
colnames(simresults)=c("totCt","p_val")

simresults=as.data.frame(simresults, stringsAsFactors = F, )
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.character)
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.numeric)

#Plot results
ggplot(simresults,aes(x=totCt,y=p_val))+geom_line(size=3)+ggtitle("Varying ALK Raw Read Count")+xlab("Count")+ylab("P-value")+theme_bw()+theme(plot.title = element_text(hjust=.5),text = element_text(size=30,face="bold"),axis.title = element_text(face="bold",size="26"),axis.text=element_text(face="bold",size="26"))
  
# ggsave("filteranalysis_Count.pdf",width = 9,height = 6,units = "in",useDingbats=F)
```
####**Varying all Filters. This runs on parallel for-loops to improve computation time.**
Since the filter cutoffs code take a long time, it is only run once. This code generates alkati_filtercutoff_allfilters.csv
```{r,eval=F}
#This chunk utilizes parallel for-loops to omptimize computational efficiency
#Run Filter Analysis using combination of filters
tic()
filename="data/all_data.csv"
alldata=read.csv(filename,sep=",",header=T,stringsAsFactors=F)

cores=detectCores()
cl= makeCluster(cores[1]-1)
registerDoParallel(cl)

#RPKM read count 1 to 100
simresults<-foreach(rawreadcount=seq(from=1,to=1500,by=10),.combine = rbind) %dopar% {
ct=1
simresults=matrix(nrow=100000,ncol=7)
library(dplyr)
#RSEM read count 1 to 1000
  for (rsem in seq(from=1,to=1000,by=10)){
    #Raw  read count 1 to 10,000
    for (rpkm in seq(from=2,to=30,by=2)) {
      
      alldata_filtered=alldata%>%
        group_by(Patid,mean_RPKM_1.19,mean_RPKM_20.29,Ratio20.29, mRNA_count,BRAF,NRAS,RSEM_normalized)%>%
        summarize(ATI=as.numeric(mRNA_count>=rawreadcount&Ratio20.29>=rpkm&RSEM_normalized>=rsem)[1])
      alldata_comp=alldata_compiler(alldata_filtered,'BRAF','NRAS','ATI','N',"N/A","N/A")[[2]]
      
      contab_pc1_genex=contab_maker(alldata_comp$Positive_Ctrl1,alldata_comp$genex,alldata_comp)
      contab_pc2_genex=contab_maker(alldata_comp$Positive_Ctrl2,alldata_comp$genex,alldata_comp)
      tot_alkati=sum(alldata_comp$genex)
      simresults[ct,1]=ct #Totalcount
      simresults[ct,2]=rpkm #RPKM
      simresults[ct,3]=rsem #RSEM
      simresults[ct,4]=rawreadcount #Alkreads
      simresults[ct,5]=fisher.test(contab_pc1_genex)$p.value #p.value
      simresults[ct,6]=fisher.test(contab_pc2_genex)$p.value #p.value
      simresults[ct,7]=tot_alkati #Total count of ALKATI patients
      ct=ct+1
    }
  }
simresults
}
stopCluster(cl)
toc()
  colnames(simresults)=c("totCt","rpkm","rsem","alkreads","p_val_pc1","p_val_pc2","tot_alk")
    
  #making simresults a dataframe
  simresults=as.data.frame(simresults, stringsAsFactors = F)
  #making everything a numeric
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.character)
simresults[colnames(simresults)] <- lapply(simresults[colnames(simresults)],as.numeric)
  
write.csv(simresults,"output/alkati_filtercutoff_allfilters.csv")
```

```{r}
simresults=read.csv("output/alkati_filtercutoff_allfilters.csv")
```

###**Figure 3a**
###**NRAS vs ALKATI**
Heat map for NRAS vs ALKATI
P-values are only shown if there were at least 20 patients that met those filters
```{r}
datapoints=simresults %>%
  filter(rsem<=500,alkreads<=1500)%>%
  group_by(rsem,alkreads,tot_alk) %>%
  summarise(min_p_val_pc2=min(p_val_pc2),rpkm_atpval=rpkm[min(p_val_pc2)][1])
datapoints[datapoints$tot_alk<20,]$min_p_val_pc2=""
datapoints$min_p_val_pc2=as.numeric(datapoints$min_p_val_pc2)

ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=min_p_val_pc2))+
      scale_fill_gradient2(low ="red" ,mid ="white" ,high ="blue",midpoint = .4 ,name="P-Value", na.value = "gray",guide = "legend",breaks=c(.1,.2,.4,.6,.8,1))+
      theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.line = element_line(colour = "black"))+
      xlab("Number of ALK Reads")+ylab("RSEM")+
      # ggtitle("Varying Combinations of Filters Does Not Create \nMutual Exclusivity Between ALK-ATI & BRAF")+
      theme_bw()+
      theme(plot.title = element_text(hjust=.5),text = element_text(size=30,face="bold"),axis.title = element_text(face="bold",size="26"),axis.text=element_text(face="bold",color="black",size="26"))
  
ggsave("output/fig2b_filtercutoff_atinras.pdf",width = 16,height = 14,units = "in",useDingbats=F)
###Regions with less than 20 patients are shaded gray.

ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=tot_alk))+
  scale_fill_gradient(,low="white",high="red",breaks = c(0,5,10,25,50,75,100),limits=c(0,100),name="ALKATI \nPatients \ndetected",guide = "legend",na.value = "red")+
  cleanup+
  theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.line = element_line(colour = "black"))+
      xlab("Number of ALK Reads")+ylab("RSEM")+
      theme_bw()+
      theme(plot.title = element_text(hjust=.5),text = element_text(size=30,face="bold"),axis.title = element_text(face="bold",size="26"),axis.text=element_text(face="bold",color="black",size="26"))

  ggsave("output/fig2b2_filtercutoff_atinras_totalalk.pdf",width = 16,height = 14,units = "in",useDingbats=F)
```

####**Heat map for NRAS vs ALKATI with count data**
```{r}
datapoints=simresults %>%
  filter(rsem<=500,alkreads<=1000)%>%
  group_by(rsem,alkreads,tot_alk) %>%
  summarise(min_p_val_pc2=min(p_val_pc2),rpkm_atpval=rpkm[min(p_val_pc2)][1])

  ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=tot_alk))+
    scale_fill_gradient2(low ="red" ,mid ="white" ,midpoint = 40,high ="blue",name="ALKATI Count")

#Trying different tiles for RPKM  
datapoints=simresults %>%
filter(rsem<=500,alkreads<=2000,rpkm%in%c(6,12,18,30))%>%
group_by(rsem,alkreads,tot_alk,rpkm) %>%
summarise(min_p_val_pc2=min(p_val_pc2))

ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=tot_alk))+facet_wrap(~rpkm)+
  scale_fill_gradient2(low ="red" ,mid ="white" ,midpoint = 40,high ="blue",name="ALKATI Count")

###Distribution of the number of ALKATI patients across different conditions  
###Remember: ALKATI patients had exon ratio of 10, rsem >100, and counts >500. i.e. green region
###The takeaway from this plot is that we can try most filters without having to worry about patient numbers. Will look at the p-values for these orange filters next
ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=tot_alk))+facet_grid(~rpkm)+
  scale_fill_gradientn(colours = rainbow(5),values=c(.1,.2,.3,.4,1),limits=c(0,100),breaks = c(9,10,20,30,40,100),name="# ALKATI \n Patients",guide = "legend")+
  ggtitle("Total ALKATI patients across exon20-29/ex1-19 ratios")+
  cleanup

###Taking a look at whether there are any statistically significant regions.
###There is only one region. We don't see mutual exclusivity at this region in the p-value plot
###only p-values between 0 and 0.1 are shown
ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=min_p_val_pc2))+facet_grid(~rpkm)+
  scale_fill_gradient2(low ="red" ,mid ="white" ,high ="blue",midpoint = .05 ,name="P-Value",guide = "legend",limits=c(0,.1))+
  cleanup

###If there are <20 patients for a set of filters, take that data out.

datapoints[datapoints$tot_alk<20,]$min_p_val_pc2=""
datapoints$min_p_val_pc2=as.numeric(datapoints$min_p_val_pc2)

###Regions with less than 20 patients are shaded gray.
ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=min_p_val_pc2))+facet_grid(~rpkm)+
  scale_fill_gradient2(low ="red" ,mid ="white" ,high ="blue",midpoint = .2 ,name="P-Value",guide = "legend",limits=c(0,1),na.value = "gray")+
  cleanup
```
##**ALKATI vs BRAF**  
###**ALKATI is not mutually exclusive with BRAF across a range of filters**
Heatmaps code continued
P-values reported are the minimum p-values across RPKM filters
Single heat map
```{r}
datapoints=simresults %>%
  filter(rsem<=500,alkreads<=1500)%>%
  group_by(rsem,alkreads,tot_alk) %>%
  summarise(min_p_val_pc1=min(p_val_pc1),rpkm_atpval=rpkm[min(p_val_pc1)][1])
datapoints[datapoints$tot_alk<20,]$min_p_val_pc1=""
datapoints$min_p_val_pc1=as.numeric(datapoints$min_p_val_pc1)

  ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=min_p_val_pc1))+
      scale_fill_gradient2(low ="red" ,mid ="white" ,high ="blue",midpoint = .4 ,name="P-Value",na.value = "gray90",guide = "legend",breaks=c(.1,.2,.4,.6,.8,1))+
    # scale_fill_gradientn(colours =c("red", "white","blue"),values=c(0,.05,1),breaks=c(0,.05,.2,1),name="P-Value",na.value = "gray",guide = "legend")+
      theme(panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            panel.background = element_blank(),
            axis.line = element_line(colour = "black"))+
      xlab("Number of ALK Reads")+ylab("RSEM")+
      # ggtitle("Varying Combinations of Filters Does Not Create \nMutual Exclusivity Between ALK-ATI & BRAF")+
      theme_bw()+
      theme(plot.title = element_text(hjust=.5),text = element_text(size=30,face="bold"),axis.title = element_text(face="bold",size="26"),axis.text=element_text(face="bold",color="black",size="26"))
  
ggsave("output/fig2b_filtercutoff_atibraf.pdf",width = 16,height = 14,units = "in",useDingbats=F)
```


Facets of heat maps across a range of ex20-29/1-19 ratios
```{r}
datapoints=simresults %>%
filter(rsem<=1000,alkreads<=2000,rpkm%in%c(6,12,18,30))%>%
group_by(rsem,alkreads,tot_alk,rpkm) %>%
summarise(min_p_val_pc1=min(p_val_pc1))

###Distribution of the number of ALKATI patients across different conditions  
###Remember: ALKATI patients had exon ratio of 10, rsem >100, and counts >500. i.e. in the green region
###The takeaway from this plot is that we can try filters in the yellow region region. Will look at the p-values for these orange filters next

ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=tot_alk))+facet_grid(~rpkm)+
  scale_fill_gradientn(colours = rainbow(5),values=c(.1,.2,.3,.4,1),limits=c(0,100),breaks = c(9,10,20,30,40,100),name="# ALKATI \n Patients",guide = "legend")+
  ggtitle("Total ALKATI patients across exon20-29/ex1-19 ratios")+
  cleanup

# P-Values Heat map
# Data is only plotted if number of ALKATI patients was >25
datapoints[datapoints$tot_alk<20,]$min_p_val_pc1=""
datapoints$min_p_val_pc1=as.numeric(datapoints$min_p_val_pc1)

ggplot(datapoints,aes(x=alkreads,y=rsem))+geom_tile(aes(fill=min_p_val_pc1))+facet_grid(~rpkm)+
  scale_fill_gradient2(low ="red" ,mid ="white" ,high ="blue",midpoint = .4 ,name="P-Value",guide = "legend",limits=c(0,1),na.value = "gray")+
  cleanup
```

