---
title: "baf3_alkati_transformations"
author: "Haider Inam"
date: "3/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..") #To change the root directory from Analysis to project
```


```{r}
library(ggsignif)
library(ggplot2)
library(knitr)
library(dplyr)
library(tictoc)
library(foreach)
library(doParallel)
source("code/alldata_compiler.R")
source("code/contab_maker.R")
######################Cleanup for GGPlot2#########################################
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))
```
##ALKATI does not seem to be sufficient to make Baf3s growth factor independent
###Displaying the population doublings of ALKATI, EML4ALK, and Vector over time
####Please not that while we performed a total of 48 ALKATI transductions, we only counted the first 9 transductions. 
```{r}
library(ggplot2)
library(reshape2)
library(Hmisc)
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

# setwd("../Box/AlkAti/figures/baf3transformationsfigure/")

baf3data=read.table("data/alkati_growthcurvedata_popdoublings.csv",header=T,stringsAsFactors = F,sep = ",")
###Removing 7th march data because the cells grew a lot in the first day post selection
baf3data=baf3data[!grepl("7-Mar",baf3data$Infection),]
baf3data=baf3data[,c(1:11)]
###If you want only eml4alk and alk ati data, do this:
# baf3data=baf3data[,c(1:8)]

#
baf3data2=melt(baf3data,id.vars = c("Time","Infection"))

baf3data2=baf3data2 %>%
  filter(value>=-10)


baf3data2=na.omit(baf3data2)

baf3data2$variable=as.character(baf3data2$variable)
###Transforming counts into actual numbers because these counts represent x1000s
# baf3data2$value=baf3data2$value*1000
###Creating a boolean for Type of infection
baf3data2$type="EML4ALK"
baf3data2$type[grep("alkati",baf3data2$variable,ignore.case = T)]="ALKATI"
baf3data2$type[grep("vector",baf3data2$variable,ignore.case = T)]="VECTOR"
#
#

ggplot(baf3data2,aes(x=Time,y=value,color=type,fill=Infection,shape=variable))+
  geom_line(size=3)+
  geom_point(size=5,shape=16)+
  # scale_y_continuous(limits = c(1,1e8))+
  ggtitle("ALK-ATI sporadically promotes growth-factor-independent proliferation")+
  xlab("Days after Transduction")+
  ylab("Number of Cell Doublings")+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
        text = element_text(size=24,face = "bold"),
        axis.title = element_text(face="bold",size="24"),
        axis.text=element_text(face="bold",size="24"))+
  theme(legend.position="none")+
  xlim(0,14)+
  scale_colour_brewer(palette="Set2")

#Can also make your own palette and color manually. Obtained from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
ggsave("output/baf3_alkati_figure_deltaadjusted_doublings.pdf",width = 16, height = 12, units = "in")

# Mean and standard deviation to grow out:
#Our criteria for cells "growing out" was when they reached two population doublings
#I extracted the number of days it took for ALKATI and EML4ALK samples to reach 2 population doublings by looking at the data of cells growing out.
head(baf3data2)
#ALKATI grew out in 9,10, and 11 days.
mean(c(9,10,11))
sd(c(9,10,11))
#Eml4ALK grew out in 6,6,6,6,6, and 5 days
mean(c(6,6,6,6,6,10))
sd(c(6,6,6,6,6,5))
```


###Converting the number of infections growing out to a barplot
```{r}
baf3_alkati_infections=data.frame(cbind(c("EML4ALK","ALKATI","VECTOR"),c(48/48*100,18/48*100,8/48*100)))
colnames(baf3_alkati_infections)=c("Infection_Type","Infection_efficiency")
baf3_alkati_infections$Infection_efficiency=as.numeric(as.character(baf3_alkati_infections$Infection_efficiency))

ggplot(baf3_alkati_infections,aes(x=Infection_Type,y=Infection_efficiency))+
  geom_col(aes(fill=Infection_Type))+
  scale_fill_brewer(palette="Set2")+
  scale_y_continuous(limits = c(0,110),name = "% Infections conferring GF-independence")+
  xlab("")+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
      text = element_text(size=24,face = "bold"),
      axis.title = element_text(face="bold",size="24"),
      axis.text=element_text(face="bold",size="24"))+
  theme(legend.position="none")+
  geom_text(label = c("48/48","18/48","8/48"),size=10,nudge_y = -10)+
  geom_signif(annotations = '*', y_position = 90 ,xmin="ALKATI", xmax="EML4ALK",size = 1,textsize = 10,nudge_y=-10)+
  geom_signif(annotations = '*', y_position = 100 ,xmin="ALKATI", xmax="VECTOR",size = 1,textsize = 10,nudge_y=-10)
ggsave("output/baf3_barplot.pdf",width = 10,height = 8,units = "in")
###Since meeting successful infection vs unsuccesfful infection are categorical variables, I will use the Chi-Sq test
#ALKATI vs EML4ALK Pvalue= 1.707e-10:
chisq.test(data.frame(rbind(c(18,30),c(48,0))))
#ALKATI vs Vector Pvalue= .038
chisq.test(data.frame(rbind(c(18,30),c(8,40))))
```

###Barplot for P24 Eliza

```{r}
#####################Barplot for P24 Elisa########################
#Inputting the lenitivirus particles that were calculated using the P24 Eliza
baf3_alkati_elisa=data.frame(cbind(c("EML4ALK","EML4ALK","EML4ALK","EML4ALK","ALKATI","ALKATI","ALKATI","ALKATI"),c("6-Nov","21-Nov","6-Jul","26-Jul","6-Nov","21-Nov","6-Jul","26-Jul"),c(590346562.5,118892812.5,46836562.5,172420312.5,744238125,410720625,46836563,75659063)))
colnames(baf3_alkati_elisa)=c("Infection_Type","Infection_Date","Lentivirus_Particles")
baf3_alkati_elisa$Lentivirus_Particles=as.numeric(as.character(baf3_alkati_elisa$Lentivirus_Particles))
baf3_alkati_elisa$Infection_Date=factor(baf3_alkati_elisa$Infection_Date,levels = (c("6-Nov","21-Nov","6-Jul","26-Jul")))
ggplot(baf3_alkati_elisa,aes(x=Infection_Date,y=Lentivirus_Particles))+
  geom_col(aes(fill=Infection_Type),position = "dodge")+
  scale_fill_brewer(palette="Set2",name="Infection\n",labels=c("ALK-ATI", "EML4ALK"))+
  ylab("Lentivirus Particles")+
  xlab("Infection Date")+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
        text = element_text(size=24,face = "bold"),
        axis.title = element_text(face="bold",size="24"),
        axis.text=element_text(face="bold",size="24",colour = "black"))+
  scale_y_continuous(trans = "log",breaks=c(1e2,1e4,1e6,1e8),labels = parse(text = c("10^2","10^4","10^6","10^8")))
ggsave("output/baf3_elisa_barplot.pdf",width = 12,height = 8,units = "in")
```

####Another way of displaying Fig 4a using unn 
```{r}
##################Baf3 figure based on raw numbers  
baf3data=read.table("data/alkati_growthcurvedata.csv",header=T,stringsAsFactors = F,sep = ",")
# ###Removing 7th march data because the cells grew a lot in the first day post selection
baf3data=baf3data[!grepl("7-Mar",baf3data$Infection),]
baf3data=baf3data[,c(1:11)]
# ###If you want only eml4alk and alk ati data, do this:
# # baf3data=baf3data[,c(1:8)]
# 
# 
baf3data2=melt(baf3data,id.vars = c("Time","Infection"))
baf3data2=na.omit(baf3data2)
baf3data2$variable=as.character(baf3data2$variable)
# ###Transforming counts into actual numbers because these counts represent x1000s
baf3data2$value=baf3data2$value*1000
# ###Creating a boolean for Type of infection
baf3data2$type="EML4ALK"
baf3data2$type[grep("alkati",baf3data2$variable,ignore.case = T)]="ALKATI"
baf3data2$type[grep("vector",baf3data2$variable,ignore.case = T)]="VECTOR"
# ###We will use the log2 ratio of the starting population to the pop of interest to get the number of doubling of a population
# ###These are called pseudo counts
# ###To avoid the problem of undefined regions when the populations go to 0, we have 
ggplot(baf3data2,aes(x=Time,y=value,color=type,fill=Infection,shape=variable))+
  geom_line(size=3)+
  geom_point(size=5,shape=16)+
  scale_y_continuous(trans = "log10")+
  ggtitle("ALK-ATI sporadically promotes growth-factor-independent proliferation")+
  xlab("Days after Transduction")+
  ylab(expression(paste(~Delta,"Cell Number (", ~10^6, ")")))+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
       text = element_text(size=24,face = "bold"),
       axis.title = element_text(face="bold",size="24"),
       axis.text=element_text(face="bold",size="24"))+
  theme(legend.position="none")
```

##ALKATI is not sufficient to rescue melanoma form a vemurafenib challenge
```{r}
melanoma_data=read.table("data/alkati_melanoma_vemurafenib_figure_data.csv",header = T,stringsAsFactors = F,sep=",")

# colnames(melanoma_data)=c("dose","skmel_pig","skmel_alkati","g361_pig","g361_alkati")
# colnames(melanoma_data)
melanoma_data2=melt(melanoma_data,id.vars = c("Dose"))
###Coercing skmel28-pig.1, .2, .3 to become the same name.
melanoma_data2$variable[grep("Skmel28.PIG",melanoma_data2$variable)]="Skmel28.PIG"
melanoma_data2$variable[grep("Skmel28.AlkAti",melanoma_data2$variable)]="Skmel28.AlkAti"
melanoma_data2$variable[grep("G361.PIG",melanoma_data2$variable)]="G361.PIG"
melanoma_data2$variable[grep("G361.AlkAti",melanoma_data2$variable)]="G361.AlkAti"

  ggplot(melanoma_data2,aes(Dose,value,color=variable))+
    stat_summary(fun.y=mean,
                 geom="point",size=10) +
    scale_x_continuous(trans = "log10") +
    stat_summary(fun.data = mean_cl_normal,
                 geom="errorbar",
                 width=.05)+
    xlab("Vemurafenib Dose (uM)") +
    ylab("Live Cell Fraction")+
    ggtitle("Dose-Response of ALK-ATI \nin Melanoma to BRAF inhibitor")+
    scale_color_manual(name="Cell Line", labels=c("\nSkmel-28 PIG","\nSkmel-28 ALKATI","\nG361 PIG","\nG361 ALKATI"),values=c("#efbc04","red","#19a4ef","#3a1aef"))+
    cleanup+
  theme(plot.title = element_text(hjust=.5),text = element_text(size=30,face="bold"),axis.title = element_text(face="bold",size="26",color="Black"),axis.text=element_text(face="bold",size="26",color="Black"))
ggsave("output/melanoma_vemurafenib_fig.pdf",width = 16, height = 12, units = "in")

###Testing whether ALKATI has a different IC50 than vector
ks.test(melanoma_data2$value[melanoma_data2$variable=="Skmel28.AlkAti"],melanoma_data2$value[melanoma_data2$variable=="Skmel28.PIG"])

ks.test(melanoma_data2$value[melanoma_data2$variable=="G361.AlkAti"],melanoma_data2$value[melanoma_data2$variable=="G361.PIG"])

#As a proof of principle showing that my test works, if I increase the dose response of G361.PIG by 25% gives my significant p.values
ks.test(melanoma_data2$value[melanoma_data2$variable=="G361.AlkAti"],melanoma_data2$value[melanoma_data2$variable=="G361.PIG"]+.25)

###P-Values: 0.99 for Skmel28, and 0.75 for G361s, showing that the ALKATI distributions were not significantly different.
```

