---
title: "baf3_alkati_transformations"
author: "Haider Inam"
date: "3/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..") #To change the root directory from Analysis to project
```


```{r}
library(reshape2)
library(ggsignif)
library(ggplot2)
library(knitr)
library(dplyr)
library(tictoc)
library(foreach)
library(doParallel)
source("code/alldata_compiler.R")
source("code/contab_maker.R")
######################Cleanup for GGPlot2#########################################
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        # axis.line = element_line(color = "black"),
        axis.text = element_text(face="bold",color="black",size="11"),
        text=element_text(size=11,face="bold"),
        axis.title=element_text(face="bold",size="11"))
# rm(list=ls())
```
##ALKATI does not seem to be sufficient to make Baf3s growth factor independent
###Displaying the population doublings of ALKATI, EML4ALK, and Vector over time
####Please not that while we performed a total of 48 ALKATI transductions, we only counted the first 9 transductions. 
```{r}
library(ggplot2)
library(reshape2)
library(Hmisc)
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))

# setwd("../Box/AlkAti/figures/baf3transformationsfigure/")

baf3data=read.table("data/alkati_growthcurvedata_popdoublings.csv",header=T,stringsAsFactors = F,sep = ",")
###Removing 7th march data because the cells grew a lot in the first day post selection
baf3data=baf3data[!grepl("7-Mar",baf3data$Infection),]
baf3data=baf3data[,c(1:11)]
###If you want only eml4alk and alk ati data, do this:
# baf3data=baf3data[,c(1:8)]

#
baf3data2=melt(baf3data,id.vars = c("Time","Infection"),variable.name ="infnum" ,value.name = "doublings")

baf3data2=baf3data2 %>%
  filter(doublings>=-10)


baf3data2=na.omit(baf3data2)

baf3data2$infnum=as.character(baf3data2$infnum)
###Transforming counts into actual numbers because these counts represent x1000s
# baf3data2$value=baf3data2$value*1000
###Creating a boolean for Type of infection
baf3data2$type="EML4ALK"
baf3data2$type[grep("alkati",baf3data2$infnum,ignore.case = T)]="ALKATI"
baf3data2$type[grep("vector",baf3data2$infnum,ignore.case = T)]="VECTOR"
#
#

ggplot(baf3data2,aes(x=Time,y=doublings,color=type,fill=Infection,shape=infnum))+
  geom_line(size=1)+
  geom_point(size=2,shape=16)+
  # scale_y_continuous(limits = c(1,1e8))+
  xlab("Days after Transduction")+
  ylab("Number of Cell Doublings")+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
        text = element_text(size=11,face = "bold"),
        axis.title = element_text(face="bold",size="11"),
        axis.text=element_text(face="bold",color="black",size="11"))+
  theme(legend.position="none")+
  xlim(0,14)+
  scale_colour_brewer(palette="Set2")

#Can also make your own palette and color manually. Obtained from http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
# ggsave("output/baf3_alkati_figure_deltaadjusted_doublings.pdf",width = 3, height = 2.25, units = "in",useDingbats=F)

# Mean and standard deviation to grow out:
#Our criteria for cells "growing out" was when they reached two population doublings
#I extracted the number of days it took for ALKATI and EML4ALK samples to reach 2 population doublings by looking at the data of cells growing out.
head(baf3data2)
#ALKATI grew out in 9,10, and 11 days.
mean(c(9,10,11))
sd(c(9,10,11))
#Eml4ALK grew out in 6,6,6,6,6, and 5 days
mean(c(6,6,6,6,6,10))
sd(c(6,6,6,6,6,5))

###Adding data for F1174 mutants that were generated on top of plvx-ALKATI-IRES-Puro using SDM
f1174data=read.table("data/alkati_growthcurvedata_popdoublings_f1174mutants.csv",header=T,stringsAsFactors = F,sep = ",")


f1174data2=melt(f1174data,id.vars = c("Time","Replicate","Condition"),value.name = "doublings",variable.name = "CellLine")

f1174data2=f1174data2 %>%
  filter(doublings>=-10)

# f1174data2=na.omit(f1174data2)

f1174data2$CellLine=as.character(f1174data2$CellLine)
f1174data2=f1174data2%>%mutate(type=case_when(
  CellLine%in%c("ALKATI")~"ALKATI",
  CellLine%in%c("F1174C","F1174V","F1174I")~"ALKATIMutant",
  CellLine%in%c("Vector")~"VECTOR",
  CellLine%in%c("EML4ALK")~"EML4ALK"
))
f1174data2=f1174data2%>%
  mutate(type_forplot=case_when(type%in%"ALKATIMutant"~"ALKATI",
                                                      TRUE~type))
ggplot(f1174data2%>%filter(!Condition%in%"Growth"),aes(x=Time/24,y=doublings,color=type_forplot,fill=CellLine,shape=factor(Replicate)))+
  geom_line(size=1)+
  geom_point(size=2,shape=16)+cleanup+
theme(plot.title = element_text(hjust=.5),
        text = element_text(size=11,face = "bold"),
        axis.title = element_text(face="bold",size="11"),
        axis.text=element_text(face="bold",color="black",size="11"))+
  scale_x_continuous(name="Time")+
  scale_y_continuous(name="Number of Cell Doublings")+
  # theme(legend.position="none")+
  # xlim(0,14)+
  scale_colour_brewer(palette="Set2")+
  xlim(0,11)+
  theme(legend.position="none")

ggsave("output/baf3_f1174_figure_deltaadjusted_doublings.pdf",width = 3, height = 2.25, units = "in",useDingbats=F)

###Combining the F1174C transformants dataframe and the Baf3 ALKATI WT transformants dataframe. Numbering the ALKATI infection numbers so that they match better with the format that I used for the BaF3 transformations last year.
f1174data2=f1174data2%>%
  mutate(infnum=CellLine,
         infnum=case_when(CellLine%in%"ALKATI"
                          &Condition%in%"IL3Sele"
                          &Replicate%in%1~"ALKATI4",
                          CellLine%in%"ALKATI"
                          &Condition%in%"IL3Sele"
                          &Replicate%in%2~"ALKATI5",
                          CellLine%in%"EML4ALK"
                          &Condition%in%"IL3Sele"
                          &Replicate%in%1~"Eml4Alk.4",
                          CellLine%in%"EML4ALK"
                          &Condition%in%"IL3Sele"
                          &Replicate%in%2~"Eml4Alk.5",
                          CellLine%in%"Vector"
                          &Condition%in%"IL3Sele"
                          &Replicate%in%1~"Vector4",
                          CellLine%in%"Vector"
                          &Condition%in%"IL3Sele"
                          &Replicate%in%2~"Vector5",
                          TRUE~paste(CellLine,Replicate,sep="")))
a=f1174data2%>%mutate(Time=Time/24)%>%dplyr::select(Time,Infection=CellLine,infnum,doublings,type,Condition)
baf3data2$Condition="IL3Sele"
brafdata_combined=rbind(baf3data2,a)

brafdata_combined=brafdata_combined%>%filter(!(Condition%in%"Growth"&Time>=9))
# c=brafdata_combined
# c=brafdata_combined%>%filter(!(Condition%in%"Growth" & Time>=6))



ggplot(brafdata_combined%>%filter(!type%in%"ALKATIMutant"),aes(x=Time,y=doublings,color=type,fill=Infection,shape=infnum,alpha=factor(Condition)))+
  geom_line(size=1)+
  geom_point(size=2,shape=16)+
  # geom_point()
  # scale_y_continuous(limits = c(1,1e8))+
  xlab("Days after Transduction")+
  ylab("Number of Cell Doublings")+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
        text = element_text(size=11,face = "bold"),
        axis.title = element_text(face="bold",size="11"),
        axis.text=element_text(face="bold",color="black",size="11"))+
  theme(legend.position="none")+
  xlim(0,13)+
  # ylim(-6,14)+
  scale_colour_brewer(palette="Set2")

# ggsave("output/baf3_alkati_figure_deltaadjusted_doublings_updated.pdf",width = 3, height = 2.25, units = "in",useDingbats=F)
```


###Converting the number of infections growing out to a barplot
```{r}
baf3_alkati_infections=data.frame(cbind(c("EML4ALK","ALKATI","VECTOR"),c(48/48*100,18/48*100,8/48*100)))
colnames(baf3_alkati_infections)=c("Infection_Type","Infection_efficiency")
baf3_alkati_infections$Infection_efficiency=as.numeric(as.character(baf3_alkati_infections$Infection_efficiency))

ggplot(baf3_alkati_infections,aes(x=Infection_Type,y=Infection_efficiency))+
  geom_col(aes(fill=Infection_Type))+
  scale_fill_brewer(palette="Set2")+
  scale_y_continuous(limits = c(0,110),name = "GF independent infections %")+
  xlab("")+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
      text = element_text(size=10,face = "bold"),
      axis.title = element_text(face="bold",size="10"),
      axis.text=element_text(face="bold",size="10",color="black"))+
  theme(legend.position="none")+
  geom_text(label = c("48/48","18/48","8/48"),size=5,nudge_y = -10)+
  geom_signif(annotations = '*', y_position = 90 ,xmin="ALKATI", xmax="EML4ALK",size = 1,textsize = 7,nudge_y=-10)+
  geom_signif(annotations = '*', y_position = 100 ,xmin="ALKATI", xmax="VECTOR",size = 1,textsize = 7,nudge_y=-10)
ggsave("output/baf3_barplot.pdf",width = 3,height = 2.25,units = "in",useDingbats=F)
###Since meeting successful infection vs unsuccesfful infection are categorical variables, I will use the Chi-Sq test
#ALKATI vs EML4ALK Pvalue= 1.707e-10:
chisq.test(data.frame(rbind(c(18,30),c(48,0))))
#ALKATI vs Vector Pvalue= .038
chisq.test(data.frame(rbind(c(18,30),c(8,40))))
```

###Barplot for P24 Eliza

```{r}
#####################Barplot for P24 Elisa########################
#Inputting the lenitivirus particles that were calculated using the P24 Eliza
baf3_alkati_elisa=data.frame(cbind(c("EML4ALK","EML4ALK","EML4ALK","EML4ALK","ALKATI","ALKATI","ALKATI","ALKATI"),c("6-Nov","21-Nov","6-Jul","26-Jul","6-Nov","21-Nov","6-Jul","26-Jul"),c(590346562.5,118892812.5,46836562.5,172420312.5,744238125,410720625,46836563,75659063)))
colnames(baf3_alkati_elisa)=c("Infection_Type","Infection_Date","Lentivirus_Particles")
baf3_alkati_elisa$Lentivirus_Particles=as.numeric(as.character(baf3_alkati_elisa$Lentivirus_Particles))
baf3_alkati_elisa$Infection_Date=factor(baf3_alkati_elisa$Infection_Date,levels = (c("6-Nov","21-Nov","6-Jul","26-Jul")))
ggplot(baf3_alkati_elisa,aes(x=Infection_Date,y=Lentivirus_Particles))+
  geom_col(aes(fill=Infection_Type),position = "dodge")+
  scale_fill_brewer(palette="Set2",name="Infection\n",labels=c("ALK-ATI", "EML4ALK"))+
  ylab("Lentivirus Particles")+
  xlab("Infection Date")+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
        text = element_text(size=10,face = "bold"),
        axis.title = element_text(face="bold",size="10"),
        axis.text=element_text(face="bold",size="10",colour = "black"))+
  scale_y_continuous(trans = "log",breaks=c(1e2,1e4,1e6,1e8),labels = parse(text = c("10^2","10^4","10^6","10^8")))
# ggsave("output/baf3_elisa_barplot.pdf",width = 4,height = 3,units = "in",useDingbats=F)
```

####Another way of displaying Fig 4a using unn 
```{r}
##################Baf3 figure based on raw numbers  
baf3data=read.table("data/alkati_growthcurvedata.csv",header=T,stringsAsFactors = F,sep = ",")
# ###Removing 7th march data because the cells grew a lot in the first day post selection
baf3data=baf3data[!grepl("7-Mar",baf3data$Infection),]
baf3data=baf3data[,c(1:11)]
# ###If you want only eml4alk and alk ati data, do this:
# # baf3data=baf3data[,c(1:8)]
# 
# 
baf3data2=melt(baf3data,id.vars = c("Time","Infection"),variable.name = "infnum",value.name = "doublings")
baf3data2=na.omit(baf3data2)
baf3data2$infnum=as.character(baf3data2$infnum)
# ###Transforming counts into actual numbers because these counts represent x1000s
baf3data2$doublings=baf3data2$doublings*1000
# ###Creating a boolean for Type of infection
baf3data2$type="EML4ALK"
baf3data2$type[grep("alkati",baf3data2$infnum,ignore.case = T)]="ALKATI"
baf3data2$type[grep("vector",baf3data2$infnum,ignore.case = T)]="VECTOR"
# ###We will use the log2 ratio of the starting population to the pop of interest to get the number of doubling of a population
# ###These are called pseudo counts
# ###To avoid the problem of undefined regions when the populations go to 0, we have 
ggplot(baf3data2,aes(x=Time,y=doublings,color=type,fill=Infection,shape=infnum))+
  geom_line(size=3)+
  geom_point(size=5,shape=16)+
  scale_y_continuous(trans = "log10")+
  ggtitle("ALK-ATI sporadically promotes growth-factor-independent proliferation")+
  xlab("Days after Transduction")+
  ylab(expression(paste(~Delta,"Cell Number (", ~10^6, ")")))+
  cleanup+
  theme(plot.title = element_text(hjust=.5),
       text = element_text(size=24,face = "bold"),
       axis.title = element_text(face="bold",size="24"),
       axis.text=element_text(face="bold",size="24"))+
  theme(legend.position="none")
```

##ALKATI is not sufficient to rescue melanoma form a vemurafenib challenge
```{r}
melanoma_data=read.table("data/alkati_melanoma_vemurafenib_figure_data.csv",header = T,stringsAsFactors = F,sep=",")
# melanoma_data=read.table("../data/alkati_melanoma_vemurafenib_figure_data.csv",header = T,stringsAsFactors = F,sep=",")

# colnames(melanoma_data)=c("dose","skmel_pig","skmel_alkati","g361_pig","g361_alkati")
# colnames(melanoma_data)
melanoma_data2=melt(melanoma_data,id.vars = c("Dose"))
###Coercing skmel28-pig.1, .2, .3 to become the same name.
melanoma_data2$variable[grep("Skmel28.PIG",melanoma_data2$variable)]="Skmel28.PIG"
melanoma_data2$variable[grep("Skmel28.AlkAti",melanoma_data2$variable)]="Skmel28.AlkAti"
melanoma_data2$variable[grep("G361.PIG",melanoma_data2$variable)]="G361.PIG"
melanoma_data2$variable[grep("G361.AlkAti",melanoma_data2$variable)]="G361.AlkAti"

ggplot(melanoma_data2,aes(Dose,value,color=variable))+
  stat_summary(fun.y=mean,
               geom="point",size=2.5) +
  scale_x_continuous(trans = "log10") +
  stat_summary(fun.data = mean_cl_normal,
               geom="errorbar",
               width=.05)+
  xlab("Vemurafenib Dose (uM)") +
  ylab("Live Cell Fraction")+
  scale_color_manual(name="Cell Line", labels=c("\nSkmel-28 PIG","\nSkmel-28 ALKATI","\nG361 PIG","\nG361 ALKATI"),values=c("#efbc04","red","#19a4ef","#3a1aef"))+
  cleanup+
theme(plot.title = element_text(hjust=.5),text = element_text(size=10,face="bold"),axis.title = element_text(face="bold",size="10",color="Black"),axis.text=element_text(face="bold",size="10",color="Black"),
      legend.position = "bottom")
# ggsave("output/melanoma_vemurafenib_fig.pdf",width = 3, height = 3, units = "in",useDingbats=F)

###Testing whether ALKATI has a different IC50 than vector
ks.test(melanoma_data2$value[melanoma_data2$variable=="Skmel28.AlkAti"],melanoma_data2$value[melanoma_data2$variable=="Skmel28.PIG"])

ks.test(melanoma_data2$value[melanoma_data2$variable=="G361.AlkAti"],melanoma_data2$value[melanoma_data2$variable=="G361.PIG"])

#As a proof of principle showing that my test works, if I increase the dose response of G361.PIG by 25% gives my significant p.values
ks.test(melanoma_data2$value[melanoma_data2$variable=="G361.AlkAti"],melanoma_data2$value[melanoma_data2$variable=="G361.PIG"]+.25)
means=melanoma_data2%>%group_by(Dose,variable)%>%dplyr::summarize(value=mean(value))
t.test(means$value[means$variable=="G361.AlkAti"],means$value[means$variable=="G361.PIG"],paired=T,alternative = "greater")
t.test(means$value[means$variable=="Skmel28.AlkAti"],means$value[means$variable=="Skmel28.PIG"],paired=T)
t.test(means$value[means$variable=="Skmel28.AlkAti"],means$value[means$variable=="G361.PIG"],paired=T)
# wilcox.test(means$value[means$variable=="G361.AlkAti"],means$value[means$variable=="G361.PIG"],paired=T,alternative = "greater")
# wilcox.test(means$value[means$variable=="Skmel28.AlkAti"],means$value[means$variable=="Skmel28.PIG"],paired=T,alternative = "greater")
###P-Values: 0.99 for Skmel28, and 0.75 for G361s, showing that the ALKATI distributions were not significantly different.

#Plotting each cell line individually
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        # axis.line = element_line(color = "black"),
        axis.text = element_text(face="bold",color="black",size="11"),
        text=element_text(size=11,face="bold"),
        axis.title=element_text(face="bold",size="11"))

melanoma_data_g361=melanoma_data2%>%filter(variable%in%c("G361.PIG","G361.AlkAti"))
ggplot(melanoma_data_g361,aes(Dose,value,color=variable))+
  stat_summary(fun.y=mean,
               geom="point",size=2.5) +
  scale_x_continuous(trans = "log10") +
  stat_summary(fun.data = mean_cl_normal,
               geom="errorbar",
               width=.05)+
  xlab("Vemurafenib Dose (uM)") +
  ylab("Live Cell Fraction")+
  scale_color_manual(name="Cell Line", labels=c("\nG361 PIG","\nG361 ALKATI"),values=c("#7B8DBF","#56B893"))+
  cleanup+
theme(plot.title = element_text(hjust=.5),text = element_text(size=10,face="bold"),axis.title = element_text(face="bold",size="10",color="Black"),axis.text=element_text(face="bold",size="10",color="Black"),
      legend.position = "none")
ggsave("output/melanoma_vemurafenib_fig_top.pdf",width = 3, height = 2.5, units = "in",useDingbats=F)

melanoma_data_skmel28=melanoma_data2%>%filter(variable%in%c("Skmel28.PIG","Skmel28.AlkAti"))
ggplot(melanoma_data_skmel28,aes(Dose,value,color=variable))+
  stat_summary(fun.y=mean,
               geom="point",size=2.5) +
  scale_x_continuous(trans = "log10") +
  stat_summary(fun.data = mean_cl_normal,
               geom="errorbar",
               width=.05)+
  xlab("Vemurafenib Dose (uM)") +
  ylab("Live Cell Fraction")+
  scale_color_manual(name="Cell Line", labels=c("\nSkmel28 PIG","\nSkmel28 ALKATI"),values=c("#7B8DBF","#56B893"))+
  cleanup+
theme(plot.title = element_text(hjust=.5),text = element_text(size=10,face="bold"),axis.title = element_text(face="bold",size="10",color="Black"),axis.text=element_text(face="bold",size="10",color="Black"),
      legend.position = "none")
ggsave("output/melanoma_vemurafenib_fig_bottom.pdf",width = 3, height = 2.5, units = "in",useDingbats=F)

```
SOS1Wt, SOS1E846K, MEKQ56P Comparisons:
```{r}
melanoma_data=read.table("data/skmel28_sos1_mekq56p_vemurafenib.csv",header = T,stringsAsFactors = F,sep=",")
melanoma_data_long=melt(melanoma_data,id.vars = "dose",na.rm = T)
melanoma_data_long$variable[grep("SOS1_E846K",melanoma_data_long$variable)]="SOS1_E846K"
melanoma_data_long$variable[grep("MEK_Q56P",melanoma_data_long$variable)]="MEK_Q56P"
melanoma_data_long$variable[grep("SOS1_WT",melanoma_data_long$variable)]="SOS1_WT"
ggplot(data = melanoma_data_long,aes(x=log(dose),y=value,color=variable))+geom_point()

ggplot(melanoma_data_long,aes(x=dose,y=value,color=variable))+geom_point(position = position_dodge(width=.01))+scale_x_continuous(trans = "log10") +cleanup+scale_colour_brewer(palette="Set2")+geom_smooth()

ggplot(melanoma_data_long,aes(dose,value,color=variable))+
    stat_summary(fun.y=mean,
               geom="point",size=2.5) +
  scale_x_continuous(trans = "log10") +
  stat_summary(fun.data = mean_cl_normal,
               geom="errorbar",
               width=.01)+
  xlab("Vemurafenib Dose (uM)") +
  ylab("Live Cell Fraction")+
  scale_color_manual(name="Cell Line", labels=c("\nSkmel-28 SOS1 E846K","\nSkmel-28 MEK Q56P","\nSkmel-28 SOS1 Wt"),values=c("#efbc04","red","#19a4ef"))+
  cleanup+
theme(plot.title = element_text(hjust=.5),text = element_text(size=10,face="bold"),axis.title = element_text(face="bold",size="10",color="Black"),axis.text=element_text(face="bold",size="10",color="Black"),
      legend.position = "bottom")

melanoma_data$sos1e846kmean=data.frame(rowMeans(melanoma_data[,c(2:5)],na.rm = T))
sos1e846kmean=data.frame(rowMeans(melanoma_data[,c(2:5)],na.rm = T))
melanoma_data$mekq56pmean=data.frame(rowMeans(melanoma_data[,c(6:9)],na.rm = T))
melanoma_data$sos1wtmean=data.frame(rowMeans(melanoma_data[,c(10:13)],na.rm = T))
sos1wtmean=data.frame(rowMeans(melanoma_data[,c(10:13)],na.rm = T))

# dose=c(3,2.69897,2.39794,2.09691,1.79588,1.49485,1.19382,1)
sos1wt=c(0.3431612,0.3592412,0.4586017,0.63938425,0.8648814,1.0462746,0.9195364,1.0000002)
sos1e846k=c(0.346128,0.421477,0.5529085,0.74714375,0.89843925,1.0010455,0.999104,1)
mekq56p=c(0.489114,0.566946,0.6174875,0.75007625,0.949546,1.02564,1.050715,1)
t.test(sos1e846k,sos1wt,paired=T,alternative = "greater")
t.test(mekq56p,sos1wt,paired=T,alternative = "greater")


```

