---
title: "pairwise_comparisons_conditional_selection_simulated_cohorts"
author: "Haider Inam"
date: "3/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..")
```

```{r,include=F}
# install.packages(c("knitr","tictoc","workflowr","VennDiagram","dplyr","foreach","doParallel","ggplot2","reshape2","RColorBrewer","devtools","ggsignif"))

library(knitr)
library(tictoc)
library(workflowr)
library(VennDiagram)
library(dplyr)
library(foreach)
library(doParallel)
library(ggplot2)
library(reshape2)
library(RColorBrewer)
library(devtools)
library(ggsignif)
source("code/contab_maker.R")
source("code/alldata_compiler.R")
source("code/quadratic_solver.R")
source("code/mut_excl_genes_generator.R")
source("code/mut_excl_genes_datapoints.R")
source("code/simresults_generator.R")

######################Cleanup for GGPlot2#########################################
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"))
```

##Part 1: Performing comparisons  between genes

### __Here, we will perform pairwise comparisons of conditional seleciton on simulated cohorts__  
### __For a detailed description of the resampling/bootstrapping algorithm for pairwise comparisons, please refer to pseudo code__  
### __For a detailed description of the algorithm for simulating genes, please refer to methods__  

###Plugging Data into simresults generator #2
These simulation have been optimized to used parallel computing.
Please note that true_OR refers to the odds ratio of the gene of interest with a positive control  
Please note that or_pair2 refers to teh odds ratio between the two positive controls (generally low)  
Note to self: just convert all the "true OR" variables to or_pair1
```{r,eval=F}
##GenePair1
cores=detectCores()
cl= makeCluster(cores[1]-1)
registerDoParallel(cl)
tic()
or_pair1=c(.01,.05,.1,.2,.3,.4,.5,.6,.7,.8,.9,1)
or_pair2=seq(.01,.1,by=.02)
incidence=seq(4,36,by=1)
cohort_size=seq(100,1000,by=200)
# unlist(mut_excl_genes_generator(500,25,.4,.01)[1])

# or_pair1=c(.01,.05,.1,.2,.3,.4,.5,.6,.7,.8,.9,1)
# or_pair2=seq(.01,.01)
# incidence=seq(4,50,by=1)
# cohort_size=seq(500,500)

# or_pair1=.8
# or_pair2=.01
# incidence=25
# cohort_size=500

true_or_vals=or_pair1
cohort_size_vals=cohort_size
gene1_total_vals=incidence

# simresults_compiled=as.data.frame(matrix(nrow=1,ncol=12)) #10 because there is an OR vals column being added at the end of the simulations
simresults_compiled=as.list(seq_len(length(or_pair2)*length(or_pair1)*length(incidence)*length(cohort_size)))

ct=1
# for(j in 1:length(gene1_total_vals)){
parfunc=foreach(j =1:length(gene1_total_vals),.combine = rbind)%dopar%{
for(l in 1:length(or_pair2)){
for(k in 1:length(cohort_size_vals)){
    for(i in 1:length(true_or_vals)){
gene_pair_1=unlist(mut_excl_genes_generator(cohort_size[k],incidence[j],or_pair1[i],or_pair2[l])[1])
gene_pair_2=unlist(mut_excl_genes_generator(cohort_size[k],incidence[j],or_pair1[i],or_pair2[l])[2])
alldata_1=mut_excl_genes_datapoints(gene_pair_1)
alldata_2=mut_excl_genes_datapoints(gene_pair_2)
alldata_comp_1=alldata_compiler(alldata_1,"gene2","gene3","gene1",'N',"N/A","N/A")[[2]]
genex_replication_prop_1=alldata_compiler(alldata_1,"gene2","gene3","gene1",'N',"N/A","N/A")[[1]]
alldata_comp_2=alldata_compiler(alldata_2,"gene2","gene3","gene1",'N',"N/A","N/A")[[2]]
genex_replication_prop_2=alldata_compiler(alldata_2,"gene2","gene3","gene1",'N',"N/A","N/A")[[1]]
#Number of simulations are the minimum of: subsampling with replacement combinations of incidence of gene pair 1, pair 2, and 10000. 10000 because I don't want to do more than 1k simulations per experiment.
nsims=min(c(choose(sum(alldata_comp_1$genex)+sum(alldata_comp_1$genex)-1,sum(alldata_comp_1$genex)),
  choose(sum(alldata_comp_2$genex)+sum(alldata_comp_2$genex)-1,sum(alldata_comp_2$genex)),1000))
# nsubsamples=sum(alldata_comp_2$genex) #Verify that this is what it is.    

# Max subsample size is whichever integer is the minimum of:
# number of genex in sample vs. number of NOTgenex or PC1 or PC2 in sample*genexreplication proportion. For example we had 165 PC1s, therefore 340-175=165 PC1s in our data. Since we want to sample 7.35% positive hits and 92.65% negative hits, a sample containing too many positive hits can impose an upper bound on our subsampling size.
alldata_comp=alldata_comp_1
genex_replication_prop=genex_replication_prop_1
maxsubsamplesize_1=min(c(genex_replication_prop*length(alldata_comp$Positive_Ctrl1),
                       floor((length(alldata_comp$Positive_Ctrl1)-sum(alldata_comp$Positive_Ctrl1))*genex_replication_prop),
                       floor((length(alldata_comp$Positive_Ctrl1)-sum(alldata_comp$Positive_Ctrl2))*genex_replication_prop),
                       floor((length(alldata_comp$Positive_Ctrl1)-sum(alldata_comp$genex))*genex_replication_prop)))
alldata_comp=alldata_comp_2
genex_replication_prop=genex_replication_prop_2
maxsubsamplesize_2=min(c(genex_replication_prop*length(alldata_comp$Positive_Ctrl1),
                       floor((length(alldata_comp$Positive_Ctrl1)-sum(alldata_comp$Positive_Ctrl1))*genex_replication_prop),
                       floor((length(alldata_comp$Positive_Ctrl1)-sum(alldata_comp$Positive_Ctrl2))*genex_replication_prop),
                       floor((length(alldata_comp$Positive_Ctrl1)-sum(alldata_comp$genex))*genex_replication_prop)))

nsubsamples=min(c(maxsubsamplesize_1,maxsubsamplesize_2)) #minimum of either of the subsamples.
nexperiments=7

#####If Cohort size is greater than gene1total, run simresults_generator for that simulation####

    if(cohort_size_vals[k]>gene1_total_vals[j]){

      simresults_pair1=simresults_generator(alldata_comp_1,7)
      simresults_pair1$gene_pair=1
      
      simresults_pair2=simresults_generator(alldata_comp_2,7)
      simresults_pair2$gene_pair=2
      
      simresults=rbind(simresults_pair1,simresults_pair2)
      simresults=as.data.frame(simresults, stringsAsFactors = F)

simresults$true_OR=true_or_vals[i]
simresults$gene1_total=gene1_total_vals[j]
simresults$cohort_size=cohort_size_vals[k]
simresults$or_pair2=or_pair2[l]

# colnames(simresults_compiled)=names(simresults)
# simresults_compiled[[i+j+k+l-3]]=simresults
simresults_compiled[[ct]]=simresults
# simresults_compiled=rbind(simresults_compiled,simresults)

    }
ct=ct+1
}
}
}
simresults_compiled
}
simresults_compiled=parfunc
simresults_compiled2=do.call(rbind,simresults_compiled)
stopCluster(cl)
toc()
# tic()
# pair1=simresults_generator(alldata_comp_1,7)
# pair1$gene_pair=1
# pair2=simresults_generator(alldata_comp_1,7)
# pair2$gene_pair=2
# aaa=rbind(pair1,pair2)
# a=c()
# a=rbind(a,aaa)
# toc()
# write.csv(simresults_compiled2,"alkati_simulations_compiled_12219.csv")
write.csv(simresults_compiled2,"data/alkati_simulations_compiled_1000_3719.csv")
# simresults_compiled[[5]]
#Looking at which conditions didn't yield a maxsimnumber of 1000 revealed that at a subsample size of 1, the max simnum was 35.
# a=simresults_compiled2%>%
#   group_by(subsample_size,exp_num,gene_pair,gene1_total,cohort_size,or_pair2)%>%
#   summarize(maxsimnum=max(sim_num))

```


##Part 2: Performing pairwise comparisons between gene pairs using KS-test  
```{r,warning=F}
simresults_compiled2=read.csv("data/alkati_simulations_compiled_1000_12319.csv",header = T)
# simresults_compiled2=read.csv("alkati_simulations_compiled_testrun_10000.csv",header = T) 
# simresults_compiled2=read.csv("alkati_simulations_compiled.csv",header = T)
simresults_stored=simresults_compiled2
simresults_compiled2=simresults_stored
simresults_compiled2=simresults_compiled2%>%
  filter((exp_num==2&gene_pair==2)|(exp_num==2&gene_pair==1))
simresults_compiled2=simresults_compiled2[,c(1,2,5,6,4,9,11,3,12,10,13,7,8)] ###This changes the simresults' columns to a more intuitive order
# simresults_compiled_concat=simresults_compiled2[,c(-1,-2,-3)]
# write.csv(simresults_compiled_concat,"alkati_simulations_compiled_1000_12319_filtered.csv")
datapoints_reshape=dcast(simresults_compiled2,exp_name+sim_num+gene1_total+subsample_size+cohort_size+true_OR+or_pair2~gene_pair,fun.aggregate = mean,value.var="p_val") 
colnames(datapoints_reshape)[8]="pair1_pval"
colnames(datapoints_reshape)[9]="pair2_pval"

###Performing KS test on the ps_p1_genex for genepair1 vs gene2
ks_results=datapoints_reshape%>%
  group_by(gene1_total,cohort_size,true_OR,or_pair2)%>%
  summarize(ks_pval=ks.test(pair1_pval,pair2_pval,alternative="less")$p.value)
  # summarize(ks_pval=ks.test(pair1_pval,pair2_pval,alternative="less")$p.value)


###Summarizing data so that min_incidence is the minimum incidence that gives a p-value for all cohorts
a=ks_results%>%
  group_by(cohort_size,true_OR,or_pair2)%>%
  summarize(sig_incidence=gene1_total[ks_pval<=.05][1])

a2=a%>%
  filter(cohort_size==500,or_pair2%in%c(.01))
##Note: these labels are only accurate for the or_pair2=.01 cohort
# label.df <- data.frame(
#   rbind(
#     # data.frame(
#     #   true_OR=c(.01),
#     #   gene1_total = c(19), 
#     #   p_val = c(1),
#     #   gene_pair=c(2)),
#     data.frame(
#       true_OR=c(.1),
#       gene1_total = c(9), 
#       p_val = c(1),
#       gene_pair=c(2)),
#     data.frame(
#       true_OR=c(1),
#       gene1_total = c(7), 
#       p_val = c(1),
#       gene_pair=c(2))))
simresults_test=simresults_compiled2%>%
  filter(cohort_size==500,or_pair2==.01,true_OR%in%c(.01,.1,1),gene1_total%in%seq(4,25,1))
simresults_test$true_OR=paste('OR of ',simresults_test$true_OR)
###Choosing parameters that most closely match ALKATI 3/6/19
simresults_test=simresults_compiled2%>%
  filter(cohort_size==300,or_pair2==.01,true_OR%in%c(.01,.1,1),gene1_total%in%seq(4,25,1))
simresults_test$true_OR=paste('OR of ',simresults_test$true_OR)

ggplot(simresults_test,aes(x=factor(gene1_total),y=log10(p_val),fill=factor(gene_pair)))+
  geom_boxplot(position = position_dodge(1))+
  facet_wrap(~true_OR,ncol=4)+
  # geom_text(data = label.df, label = "*",size=15,color="black")+
  scale_fill_brewer(palette = "Set2",name="Gene Pair",labels=c("GOI", "Positive Control"),direction = -1)+
  cleanup+
  scale_y_continuous(name="Log(P-Value from Fisher's Test)",
                     limits = c(-3,0))+
  scale_x_discrete(name="Incidence of GOI in simulated cohort",limits=c("5","10","15","20","25"))+
  theme(plot.title = element_text(hjust=.5),
        text = element_text(size=26,face="bold"),
        axis.title = element_text(face="bold",size="26",color="black"),
        axis.text=element_text(face="bold",size="24",color="black"))
    
# ggsave("output/fig1c_3719_4.pdf",height = 8,width = 20,units = "in")

# +xlab("Subsample Size")+ylab("P-value")+theme_bw()+theme(plot.title = element_text(hjust=.5),text = element_text(size=26,face="bold"),axis.title = element_text(face="bold",size="26"),axis.text=element_text(face="bold",size="26"))
# +scale_color_discrete(name="Pairwise \nComparison\n",labels=c("OR 0.1 to 1\n","OR 0.1 to 0.01"))

# If you wanna look at true_OR on the x-axis:
# simresults_test=simresults_compiled2%>%
#   filter(cohort_size==500,or_pair2==.01,gene1_total%in%c(10))
# ggplot(simresults_test,aes(x=factor(true_OR),y=p_val,fill=factor(gene_pair)))+
#   geom_boxplot(position = position_dodge(1))+
#   facet_wrap(~factor(gene1_total))

```


##Part 3: Data Analysis and Plotting  
