---
title: "updated_resampling_strategy"
author: "Haider Inam"
date: "5/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=rprojroot::find_rstudio_root_file())
```

```{r}
library(knitr)
library(tictoc)
library(workflowr)
library(VennDiagram)
library(dplyr)
library(foreach)
library(doParallel)
library(ggplot2)
library(reshape2)
library(RColorBrewer)
library(devtools)
library(ggsignif)
library(plotly)
library(BiocManager)
library(drc)
# library(ensembldb) #Loading this with Dplyr commands seems to throw an error in Rmd
# library(EnsDb.Hsapiens.v86) #Loading this with Dplyr commands seems to throw an error in Rmd

# source("code/contab_maker.R")
# source("code/alldata_compiler.R")
# source("code/quadratic_solver.R")
# source("code/mut_excl_genes_generator.R")
# source("code/mut_excl_genes_datapoints.R")
# source("code/simresults_generator.R")

source("../code/contab_maker.R")
source("../code/alldata_compiler.R")
source("../code/quadratic_solver.R")
source("../code/mut_excl_genes_generator.R")
source("../code/mut_excl_genes_datapoints.R")
source("../code/simresults_generator.R")


######################Cleanup for GGPlot2#########################################
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(color = "black"),
        axis.text = element_text(face="bold",color="black",size="11"),
        text=element_text(size=11,face="bold"),
        axis.title=element_text(face="bold",size="11"))
```


```{r}
  nameposctrl1<-'BRAF'
  #Positive control 1
  nameposctrl2<-'NRAS'
  #Oncogene in Question
  namegene<-'ATI'
  #Mutation Boolean (Y or N)
  mtn<-'N'
  #Name Mutation for Positive Ctrl 1
  nameposctrl1mt<-'V600E'
  #Name of Mutation for Positive Ctrl 2
  nameposctrl2mt<-'Q61L'

# alldata=read.csv("output/all_data_skcm.csv",sep=",",header=T,stringsAsFactors=F)
alldata=read.csv("../output/all_data_skcm.csv",sep=",",header=T,stringsAsFactors=F)
head(alldata)
```


```{r}
###Not mutation specific generation of counts###
alldata_comp=alldata_compiler(alldata,nameposctrl1,nameposctrl2,namegene,'N',"N/A","N/A")[[2]]
head(alldata_comp)
```
```{r}
sum(alldata_comp$Positive_Ctrl2)/length(alldata_comp$Positive_Ctrl1)
#subsamplesize here is defined as the number of hits of the respective gene, the PC1 or PC2 in this case.
#How do you downsample the data to a sss of PC1 or genex: the sample command takes in how many rows you wanna randomly sample. For example, there are 96 NRAS hits in the entire dataset and is at a frequency of 27%. So if you wanted it to be 40 instead, you would have to subsample 40/0.27=148 rows of data
###Probably wanna initialize the resampled dataframe as a preallocated matrix
# resampled_pc1_cum=matrix(NA,1,7)
# resampled_pc1_cum=resampled_pc1[1,]
pc1pc2_pval_all=matrix(NA,1,2)
pc1goi_pval_all=matrix(NA,1,2)
for(sss in seq(5,50,by=3)){
  pc1pc2_pval_cum=NA #of course don't wanna use this number always
  pc1goi_pval_cum=NA #of course don't wanna use this number always

  for(i in 1:100){
    # i=1
    resampled_pc1=alldata_comp[sample(length(alldata_comp$Positive_Ctrl1),sss*length(alldata_comp$Positive_Ctrl1)/sum(alldata_comp$Positive_Ctrl1),replace = T,prob = NULL),]
    
    resampled_goi=alldata_comp[sample(length(alldata_comp$Positive_Ctrl1),sss*length(alldata_comp$Positive_Ctrl1)/sum(alldata_comp$genex),replace = T,prob = NULL),]
  
    con_tab_sample_pctrl1_pctrl2=contab_maker(resampled_pc1$Positive_Ctrl1,resampled_pc1$Positive_Ctrl2,resampled_pc1)
    con_tab_sample_pctrl1_goi=contab_maker(resampled_goi$Positive_Ctrl1,resampled_goi$genex,resampled_goi)
    p_s_p1_p2=fisher.test(con_tab_sample_pctrl1_pctrl2)
    p_s_p1_goi=fisher.test(con_tab_sample_pctrl1_goi)
    
    pc1pc2_pval=p_s_p1_p2$p.value
    pc1goi_pval=p_s_p1_goi$p.value
    pc1pc2_pval_cum=c(pc1pc2_pval_cum,pc1pc2_pval)
    pc1goi_pval_cum=c(pc1goi_pval_cum,pc1goi_pval)
    
    # conttab_pc1=resampled_pc1_cum=rbind(resampled_pc1,resampled_pc1_cum)
  }
  pc1pc2_pval_cum=na.omit(pc1pc2_pval_cum)
  pc1goi_pval_cum=na.omit(pc1goi_pval_cum)
  # class(pc1pc2_pval_cum)
  pc1pc2_pval_cum=cbind(pc1pc2_pval_cum,sss)
  pc1goi_pval_cum=cbind(pc1goi_pval_cum,sss)
  # pc1pc2_pval_cum=as.matrix(pc1pc2_pval_cum)
  pc1pc2_pval_all=rbind(pc1pc2_pval_all,pc1pc2_pval_cum)
  pc1goi_pval_all=rbind(pc1goi_pval_all,pc1goi_pval_cum)
}
pc1pc2_pval_all=na.omit(pc1pc2_pval_all)
pc1pc2_pval_all=as.data.frame(pc1pc2_pval_all)
colnames(pc1pc2_pval_all)=c("pval","sss")
pc1pc2_pval_all$experiment="pc1pc2"

pc1goi_pval_all=na.omit(pc1goi_pval_all)
pc1goi_pval_all=as.data.frame(pc1goi_pval_all)
colnames(pc1goi_pval_all)=c("pval","sss")
pc1goi_pval_all$experiment="pc1goi"
simulations_df_comb=rbind(pc1goi_pval_all,pc1pc2_pval_all)
# for each iteration of the bootstrap, put resampled data in a cont table. Could probably do that outside of the main for-loop for faster computation
# class(pc1pc2_pval_all$sss)
ggplot(pc1pc2_pval_all%>%filter(),aes(x=factor(sss),y=pval))+geom_boxplot()+geom_hline(yintercept=0.3,linetype="dashed",color="red")
ggplot(pc1pc2_pval_all%>%filter(),aes(x=factor(sss),y=pval))+geom_boxplot()+geom_hline(yintercept=0.3,linetype="dashed",color="red")+scale_y_continuous(trans="log10",name="P-value distribution of BRAF vs NRAS")+scale_x_discrete("Subsample Size")+cleanup

ggplot(simulations_df_comb,aes(x=factor(sss),y=pval,fill=experiment))+geom_boxplot()+geom_hline(yintercept=0.3,linetype="dashed",color="red")+scale_y_continuous(trans="log10",name="P-value distribution of BRAF vs NRAS")+scale_x_discrete("Subsample Size")+cleanup+scale_fill_manual(values=c("#FFD92F","#E78AC3"))

# scale_color_manual(values =c("#FFD92F","#E78AC3","#8DA0CB"),name="Mutation")

```

####I coded the following part up for David's idea of simple reasmpling but realized that doing simple resamping gives us a downsampled abundance but preserves the frequency of the positive controls. Therefore, I'm gonna use out simresults_generator as-is without the max-subsamplesize clause and the simresults clause.

```{r}
sum(alldata_comp$Positive_Ctrl2)/length(alldata_comp$Positive_Ctrl1)
#subsamplesize here is defined as the number of hits of the respective gene, the PC1 or PC2 in this case.
#How do you downsample the data to a sss of PC1 or genex: the sample command takes in how many rows you wanna randomly sample. For example, there are 96 NRAS hits in the entire dataset and is at a frequency of 27%. So if you wanted it to be 40 instead, you would have to subsample 40/0.27=148 rows of data
###Probably wanna initialize the resampled dataframe as a preallocated matrix
# resampled_pc1_cum=matrix(NA,1,7)
# resampled_pc1_cum=resampled_pc1[1,]
pc1pc2_pval_all=matrix(NA,1,2)
pc1goi_pval_all=matrix(NA,1,2)
for(sss in seq(5,50,by=3)){
  pc1pc2_pval_cum=NA #of course don't wanna use this number always
  pc1goi_pval_cum=NA #of course don't wanna use this number always

  for(i in 1:100){
    # i=1
    resampled_pc1=alldata_comp[sample(length(alldata_comp$Positive_Ctrl1),sss*length(alldata_comp$Positive_Ctrl1)/sum(alldata_comp$Positive_Ctrl1),replace = T,prob = NULL),]
    
    resampled_goi=alldata_comp[sample(length(alldata_comp$Positive_Ctrl1),sss*length(alldata_comp$Positive_Ctrl1)/sum(alldata_comp$genex),replace = T,prob = NULL),]
  
    con_tab_sample_pctrl1_pctrl2=contab_maker(resampled_pc1$Positive_Ctrl1,resampled_pc1$Positive_Ctrl2,resampled_pc1)
    con_tab_sample_pctrl1_goi=contab_maker(resampled_goi$Positive_Ctrl1,resampled_goi$genex,resampled_goi)
    p_s_p1_p2=fisher.test(con_tab_sample_pctrl1_pctrl2)
    p_s_p1_goi=fisher.test(con_tab_sample_pctrl1_goi)
    
    pc1pc2_pval=p_s_p1_p2$p.value
    pc1goi_pval=p_s_p1_goi$p.value
    pc1pc2_pval_cum=c(pc1pc2_pval_cum,pc1pc2_pval)
    pc1goi_pval_cum=c(pc1goi_pval_cum,pc1goi_pval)
    
    # conttab_pc1=resampled_pc1_cum=rbind(resampled_pc1,resampled_pc1_cum)
  }
  pc1pc2_pval_cum=na.omit(pc1pc2_pval_cum)
  pc1goi_pval_cum=na.omit(pc1goi_pval_cum)
  # class(pc1pc2_pval_cum)
  pc1pc2_pval_cum=cbind(pc1pc2_pval_cum,sss)
  pc1goi_pval_cum=cbind(pc1goi_pval_cum,sss)
  # pc1pc2_pval_cum=as.matrix(pc1pc2_pval_cum)
  pc1pc2_pval_all=rbind(pc1pc2_pval_all,pc1pc2_pval_cum)
  pc1goi_pval_all=rbind(pc1goi_pval_all,pc1goi_pval_cum)
}
pc1pc2_pval_all=na.omit(pc1pc2_pval_all)
pc1pc2_pval_all=as.data.frame(pc1pc2_pval_all)
colnames(pc1pc2_pval_all)=c("pval","sss")
pc1pc2_pval_all$experiment="pc1pc2"

pc1goi_pval_all=na.omit(pc1goi_pval_all)
pc1goi_pval_all=as.data.frame(pc1goi_pval_all)
colnames(pc1goi_pval_all)=c("pval","sss")
pc1goi_pval_all$experiment="pc1goi"
simulations_df_comb=rbind(pc1goi_pval_all,pc1pc2_pval_all)
# for each iteration of the bootstrap, put resampled data in a cont table. Could probably do that outside of the main for-loop for faster computation
# class(pc1pc2_pval_all$sss)
ggplot(pc1pc2_pval_all%>%filter(),aes(x=factor(sss),y=pval))+geom_boxplot()+geom_hline(yintercept=0.3,linetype="dashed",color="red")
ggplot(pc1pc2_pval_all%>%filter(),aes(x=factor(sss),y=pval))+geom_boxplot()+geom_hline(yintercept=0.3,linetype="dashed",color="red")+scale_y_continuous(trans="log10",name="P-value distribution of BRAF vs NRAS")+scale_x_discrete("Subsample Size")+cleanup

ggplot(simulations_df_comb,aes(x=factor(sss),y=pval,fill=experiment))+geom_boxplot()+geom_hline(yintercept=0.3,linetype="dashed",color="red")+scale_y_continuous(trans="log10",name="P-value distribution of BRAF vs NRAS")+scale_x_discrete("Subsample Size")+cleanup+scale_fill_manual(values=c("#FFD92F","#E78AC3"))

# scale_color_manual(values =c("#FFD92F","#E78AC3","#8DA0CB"),name="Mutation")

```
Running pairwise comparisons for ALKATI
```{r}
nsims=100
data=simresults_generator(alldata_comp,7,100,20,.11)
# simresults_data_comp=matrix(NA,1,7)
simresults_data_comp=data[1,]
for(nsubsamples in seq(3,20,3)){
  data=simresults_generator(alldata_comp,7,200,nsubsamples,.11)
  simresults_data_comp=rbind(simresults_data_comp,data)
}
simresults_data_comp=simresults_data_comp[-1,]

ggplot(simresults_data_comp%>%filter(exp_name%in%c("p_s_p1_genex","p_s_p1_p2")),aes(x=factor(subsample_size),y=p_val,fill=exp_name))+geom_boxplot()
```

Running pairwise comparisons for simulated dataset
```{r}
# i=0.9 #goi control gene pair OR
# k=500 #cohort size
# l=0.08 #positive control gene pair OR

gene_pair_1=unlist(mut_excl_genes_generator(500,20,0.9,0.08)[1])
gene_pair_2=unlist(mut_excl_genes_generator(500,20,0.9,0.08)[2])
alldata_1=mut_excl_genes_datapoints(gene_pair_1)
alldata_2=mut_excl_genes_datapoints(gene_pair_2)
alldata_comp_1=alldata_compiler(alldata_1,"gene2","gene3","gene1",'N',"N/A","N/A")[[2]]
genex_replication_prop_1=alldata_compiler(alldata_1,"gene2","gene3","gene1",'N',"N/A","N/A")[[1]]
alldata_comp_2=alldata_compiler(alldata_2,"gene2","gene3","gene1",'N',"N/A","N/A")[[2]]
genex_replication_prop_2=alldata_compiler(alldata_2,"gene2","gene3","gene1",'N',"N/A","N/A")[[1]]

nsubsamples=10
nsims=100
simresults_pair1=simresults_generator(alldata_comp_1,7,nsims,nsubsamples,genex_replication_prop_1)
simresults_pair1$gene_pair=1
      
simresults_pair2=simresults_generator(alldata_comp_2,7,nsims,nsubsamples,genex_replication_prop_2)
simresults_pair2$gene_pair=2
```

```{r}

```

