---
title: "updated_resampling_strategy2"
author: "Haider Inam"
date: "5/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir=rprojroot::find_rstudio_root_file())
```

```{r,include=F}
# rm(list=ls())
#Loading required packages
library(knitr)
library(tictoc)
library(workflowr)
library(VennDiagram)
library(dplyr)
library(foreach)
library(doParallel)
library(ggplot2)
library(reshape2)
library(RColorBrewer)
library(devtools)
library(ggsignif)
library(plotly)
library(BiocManager)
######################Cleanup for GGPlot2#########################################
cleanup=theme_bw() +
  theme(plot.title = element_text(hjust=.5),
        panel.grid.major = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.background = element_blank(),
        # axis.line = element_line(color = "black"),
        axis.text = element_text(face="bold",color="black",size="11"),
        text=element_text(size=11,face="bold"),
        axis.title=element_text(face="bold",size="11"))
```

```{r}

# library(ensembldb) #Loading this with Dplyr commands seems to throw an error in Rmd
# library(EnsDb.Hsapiens.v86) #Loading this with Dplyr commands seems to throw an error in Rmd
# source("../code/mut_excl_genes_generator3.R")
# source("../code/mut_excl_genes_datapoints.R")

source("code/contab_maker.R")
source("code/contab_simulator.R")
source("code/contab_downsampler.R")
source("code/alldata_compiler.R")
source("code/mut_excl_genes_generator.R")

# source("../code/contab_maker.R")
# source("../code/contab_simulator.R")
# source("../code/contab_downsampler.R")
# source("../code/alldata_compiler.R")
# source("../code/mut_excl_genes_generator.R")


```

```{r,include=F}
# source("code/mut_excl_genes_generator.R")
# source("code/archive/mut_excl_genes_generator3.R")
# source("code/archive/mut_excl_genes_generator_old.R")
# mut_excl_genes_generator(100,12,.01,.1)
# mut_excl_genes_generator(100,12,.1,.01)
```

  
### Pairwise Comparisons using Simulated Genes###  
  
#### This is the part of the code where I performed pairwise comparisons on gene pairs having different abundances, odds ratios, and cohort sizes.
```{r}

# rm(list=ls())
or_pair1=c(.01,.05,.1,.2,.3,.4,.5,.6,.7,.8,.9,1)
or_pair2=seq(.01,.2,by=.02)
incidence=seq(4,36,by=2)
cohort_size=seq(100,1000,by=200)

# or_pair1=c(.01,.05)
# or_pair2=seq(.01,.1,by=.05)
# incidence=seq(4,36,by=20)
# cohort_size=seq(100,1000,by=400)

    # or_pair1=.05
    # or_pair2=.01
    # incidence=12
    # cohort_size=500
    # i=1
    # j=1
    # k=1
    # l=1

true_or_vals=or_pair1
cohort_size_vals=cohort_size
gene1_total_vals=incidence

simresults_compiled_alldata=as.list(length(or_pair2)*length(or_pair1)*length(incidence)*length(cohort_size))
simresults_compiled=matrix(,length(or_pair2)*length(or_pair1)*length(incidence)*length(cohort_size),ncol=16)
ct=1
for(j in 1:length(incidence)){
  tic()
for(l in 1:length(or_pair2)){
  for(k in 1:length(cohort_size_vals)){
    for(i in 1:length(true_or_vals)){
gene_pair_1=unlist(mut_excl_genes_generator(cohort_size[k],incidence[j],or_pair1[i],or_pair2[l])[1])
gene_pair_1_table=rbind(c(gene_pair_1[1],gene_pair_1[2]),c(gene_pair_1[3],gene_pair_1[4])) ###make sure these are the right indices
gene_pair_2=unlist(mut_excl_genes_generator(cohort_size[k],incidence[j],or_pair1[i],or_pair2[l])[2])
gene_pair_2_table=rbind(c(gene_pair_2[1],gene_pair_2[2]),c(gene_pair_2[3],gene_pair_2[4]))
# alldata_1=mut_excl_genes_datapoints(gene_pair_1)
# 
# alldata_2=mut_excl_genes_datapoints(gene_pair_2)
# alldata_comp_1=alldata_compiler(alldata_1,"gene2","gene3","gene1",'N',"N/A","N/A")[[2]]
# 
# genex_replication_prop_1=alldata_compiler(alldata_1,"gene2","gene3","gene1",'N',"N/A","N/A")[[1]]
# alldata_comp_2=alldata_compiler(alldata_2,"gene2","gene3","gene1",'N',"N/A","N/A")[[2]]
# genex_replication_prop_2=alldata_compiler(alldata_2,"gene2","gene3","gene1",'N',"N/A","N/A")[[1]]


###Calculating Odds ratios and GOI frequencies for the raw data###
# cohort_size_curr=length(alldata_comp$Positive_Ctrl1)
cohort_size_curr=cohort_size[k]
# pc1pc2_contab_counts=contab_maker(alldata_comp$Positive_Ctrl1,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
pc1pc2_contab_counts=gene_pair_2_table
# goipc1_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl1,alldata_comp)[2:1, 2:1]
goipc1_contab_counts=gene_pair_1_table


# goipc2_contab_counts=contab_maker(alldata_comp$genex,alldata_comp$Positive_Ctrl2,alldata_comp)[2:1, 2:1]
# pc1pc2_contab_probabilities=pc1pc2_contab_counts/cohort_size_curr
# goipc1_contab_probabilities=goipc1_contab_counts/cohort_size_curr
pc1pc2_contab_probabilities=pc1pc2_contab_counts
goipc1_contab_probabilities=goipc1_contab_counts
# goipc2_contab_probabilities=goipc2_contab_counts/cohort_size

or_pc1pc2=pc1pc2_contab_probabilities[1,1]*pc1pc2_contab_probabilities[2,2]/(pc1pc2_contab_probabilities[1,2]*pc1pc2_contab_probabilities[2,1])
or_goipc1=goipc1_contab_probabilities[1,1]*goipc1_contab_probabilities[2,2]/(goipc1_contab_probabilities[1,2]*goipc1_contab_probabilities[2,1])
# or_goipc2=goipc2_contab_probabilities[1,1]*goipc2_contab_probabilities[2,2]/(goipc2_contab_probabilities[1,2]*goipc2_contab_probabilities[2,1])

goi_freq=goipc1_contab_probabilities[1,1]+goipc1_contab_probabilities[1,2]
# goi_freq=.01
# class(goi_freq)

###

###Downsampling PC1 to the probability of GOI without changing ORs###
###The function below converts contingency table data to a new contingency table in which the data is downsampled to the desired frequency, aka the frequency of the GOI in this case###
pc1new_pc2_contab=contab_downsampler(pc1pc2_contab_probabilities,goi_freq)
goinew_pc1_contab=contab_downsampler(goipc1_contab_probabilities,goi_freq)
# goinew_pc2_contab=contab_downsampler(goipc2_contab_probabilities,goi_freq)
###original contab:
# head(pc1pc2_contab_probabilities) 
###downsampled contab:
# head(pc1new_pc2_contab) 

pc1rawpc2_contabs_sims=contab_simulator(pc1pc2_contab_probabilities,1000,cohort_size_curr)
pc1pc2_contabs_sims=contab_simulator(pc1new_pc2_contab,1000,cohort_size_curr)
goipc1_contabs_sims=contab_simulator(goinew_pc1_contab,1000,cohort_size_curr)
# goipc2_contabs_sims=contab_simulator(goinew_pc2_contab,1000,cohort_size)
# head(pc1pc2_contabs_sims) #each row in this dataset is a new contab
pc1rawpc2_contabs_sims=data.frame(pc1rawpc2_contabs_sims)
  pc1rawpc2_contabs_sims=pc1rawpc2_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))

pc1pc2_contabs_sims=data.frame(pc1pc2_contabs_sims)
  pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))
  
goipc1_contabs_sims=data.frame(goipc1_contabs_sims)
  goipc1_contabs_sims=goipc1_contabs_sims%>%
  mutate(or=p11*p00/(p10*p01))
  

  
# goipc2_contabs_sims=data.frame(goipc2_contabs_sims)
#   goipc2_contabs_sims=goipc2_contabs_sims%>%
#   mutate(or=p11*p00/(p10*p01))
pc1rawpc2_contabs_sims$comparison="pc1rawpc2"
pc1pc2_contabs_sims$comparison="pc1pc2"
goipc1_contabs_sims$comparison="goipc1"
# goipc2_contabs_sims$comparison="goipc2"
or_median_raw=quantile(pc1rawpc2_contabs_sims$or,na.rm = T)[3]
or_uq_raw=quantile(pc1rawpc2_contabs_sims$or,na.rm = T)[4]
or_median_downsampled=quantile(pc1pc2_contabs_sims$or,na.rm = T)[3]
or_uq_downsampled=quantile(pc1pc2_contabs_sims$or,na.rm = T)[4]



pc1rawpc2_contabs_sims=pc1rawpc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>=or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>=or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
goipc1_contabs_sims=goipc1_contabs_sims%>%
  mutate(isgreater_raw_median=case_when(or>or_median_raw~1,
                             TRUE~0),
         isgreater_raw_uq=case_when(or>or_uq_raw~1,
                             TRUE~0),
         isgreater_median=case_when(or>or_median_downsampled~1,
                             TRUE~0),
         isgreater_uq=case_when(or>or_uq_downsampled~1,
                             TRUE~0)
         )
# pc1pc2_contabs_sims=pc1pc2_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
# goipc1_contabs_sims=goipc1_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
# goipc2_contabs_sims=goipc2_contabs_sims%>%
#   mutate(isgreater=case_when(or>=or_pc1pc2~1,
#                              TRUE~0))
pc1rawpc2_isgreater_raw_median=sum(pc1rawpc2_contabs_sims$isgreater_raw_median)
pc1rawpc2_isgreater_raw_uq=sum(pc1rawpc2_contabs_sims$isgreater_raw_uq)
pc1rawpc2_isgreater_median=sum(pc1rawpc2_contabs_sims$isgreater_median)
pc1rawpc2_isgreater_uq=sum(pc1rawpc2_contabs_sims$isgreater_uq)

pc1pc2_isgreater_raw_median=sum(pc1pc2_contabs_sims$isgreater_raw_median)
pc1pc2_isgreater_raw_uq=sum(pc1pc2_contabs_sims$isgreater_raw_uq)
pc1pc2_isgreater_median=sum(pc1pc2_contabs_sims$isgreater_median)
pc1pc2_isgreater_uq=sum(pc1pc2_contabs_sims$isgreater_uq)

goipc1_isgreater_raw_median=sum(goipc1_contabs_sims$isgreater_raw_median)
goipc1_isgreater_raw_uq=sum(goipc1_contabs_sims$isgreater_raw_uq)
goipc1_isgreater_median=sum(goipc1_contabs_sims$isgreater_median)
goipc1_isgreater_uq=sum(goipc1_contabs_sims$isgreater_uq)


# pc1rawpc2_isgreater=sum(pc1rawpc2_contabs_sims$isgreater)
# pc1pc2_isgreater=sum(pc1pc2_contabs_sims$isgreater)
# goipc1_isgreater=sum(goipc1_contabs_sims$isgreater)

simresults=c(cohort_size[k],
             incidence[j],
             or_pair1[i],
             or_pair2[l],
             pc1rawpc2_isgreater_raw_median,
             pc1rawpc2_isgreater_raw_uq,
             pc1rawpc2_isgreater_median,
             pc1rawpc2_isgreater_uq,
             pc1pc2_isgreater_raw_median,
             pc1pc2_isgreater_raw_uq,
             pc1pc2_isgreater_median,
             pc1pc2_isgreater_uq,
             goipc1_isgreater_raw_median,
             goipc1_isgreater_raw_uq,
             goipc1_isgreater_median,
             goipc1_isgreater_uq)

simresults_alldata=c(cohort_size[k],
             incidence[j],
             or_pair1[i],
             or_pair2[l],
    list(goipc1_contabs_sims$or),
    list(pc1pc2_contabs_sims$or))

######Comment out the following section if you wish to not gather all the 1,000 Odds ratios for each simulation####
simresults_compiled_alldata[[ct]]=simresults_alldata
# a=simresults_compiled_alldata[[1]]

simresults_compiled[ct,]=simresults

ct=ct+1
    }
  }
}
  toc()
}

```

```{r}
simresults_compiled=data.frame(simresults_compiled)
colnames(simresults_compiled)=c("cohort_size",
                                "incidence",
                                "or1",
                                "or2",
                                "pc1rawpc2_isgreater_raw_median",
                                "pc1rawpc2_isgreater_raw_uq",
                                "pc1rawpc2_isgreater_median",
                                "pc1rawpc2_isgreater_uq",
                                "pc1pc2_isgreater_raw_median",
                                "pc1pc2_isgreater_raw_uq",
                                "pc1pc2_isgreater_median",
                                "pc1pc2_isgreater_uq",
                                "goipc1_isgreater_raw_median",
                                "goipc1_isgreater_raw_uq",
                                "goipc1_isgreater_median",
                                "goipc1_isgreater_uq")
simresults_compiled$delta_median=simresults_compiled$goipc1_isgreater_raw_median-simresults_compiled$goipc1_isgreater_median
simresults_compiled$delta_uq=simresults_compiled$goipc1_isgreater_raw_uq-simresults_compiled$goipc1_isgreater_uq

simresults_compiled=simresults_compiled%>%
  mutate(fp_corrected_95=case_when(
    goipc1_isgreater_raw_uq<950~-1,
     (goipc1_isgreater_raw_uq>=950&(goipc1_isgreater_raw_uq-delta_uq)<=950)~1,
    TRUE~0),
    fp_corrected_99=case_when(
      goipc1_isgreater_raw_uq<990~-1,
    (goipc1_isgreater_raw_uq>=990&(goipc1_isgreater_raw_uq-delta_uq)<=990)~1,
    TRUE~0))

# simresults_compiled$goipc1_isgreater_percent=simresults_compiled$goipc1_isgreater*100/1000
simresults_subset=simresults_compiled%>%filter(cohort_size%in%500)

```
  
### Here I am plotting the score obtained for all simulations. The score is defined as the percentage of GOI vs PC1 trials that fall in the median of the odds ratios of the frequency corrected PC1 vs PC2 simulations.
```{r}
ggplot(simresults_compiled%>%filter(or1<=.4),aes(x=incidence,y=or2))+geom_tile(aes(fill=goipc1_isgreater_median/10))+facet_grid(cohort_size~or1)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")
```
  
### Here I am plotting the score obtained for all simulations. The score is defined as the percentage of GOI vs PC1 trials that fall in the upper quartile of the odds ratios of the frequency corrected PC1 vs PC2 simulations. i.e. it is a less stringent test###
```{r}
ggplot(simresults_compiled%>%filter(or1<=.4),aes(x=incidence,y=or2))+geom_tile(aes(fill=goipc1_isgreater_uq/10))+facet_grid(cohort_size~or1)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Percent succesful trials")
```
  
### A few more ways to look at the plots above:
```{r}

ggplot(simresults_compiled%>%filter(or1<=.4,cohort_size%in%500),aes(x=incidence,y=or2))+geom_tile(aes(fill=goipc1_isgreater_median/10))+facet_wrap(~or1)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")

ggplot(simresults_compiled%>%filter(or1<=.4,cohort_size%in%500),aes(x=incidence,y=or2))+geom_tile(aes(fill=goipc1_isgreater_median/10))+facet_wrap(~or1,ncol=6)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")

ggplot(simresults_compiled%>%filter(or1%in%c(.05,0.1,.5),cohort_size%in%500),aes(x=incidence,y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_median/10))+facet_wrap(~or1,ncol=6)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_continuous(expand = c(0,0),name="Incidence")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1vsPC2")+ggtitle("Odds ratio of GOIvsPC1")+
  theme(plot.title = element_text(hjust = 0.5))

# ggsave("score_heatmap.pdf",width=8,heigh=3,units="in",useDingbats=F)

ggplot(simresults_compiled%>%filter(or1%in%c(.05,0.1,.5),cohort_size%in%c(100,500,900)),aes(x=incidence,y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_median/10))+facet_grid(cohort_size~or1)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_continuous(expand = c(0,0),name="Incidence")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1vsPC2")+ggtitle("Odds ratio of GOIvsPC1")+
  theme(plot.title = element_text(hjust = 0.5))

# ggsave("score_heatmap_supplement.pdf",width=8,heigh=7,units="in",useDingbats=F)


ggplot(simresults_compiled%>%filter(or2%in%.05,cohort_size%in%c(500)),aes(x=factor(or1),y=goipc1_isgreater_median/10))+geom_boxplot()+scale_x_discrete(name="Odds ratio of PC1vsPC2")+
  scale_y_continuous(name="Score")+
  theme(plot.title = element_text(hjust = 0.5))+cleanup

# ggsave("score_plot.pdf",width=8,heigh=3,units="in",useDingbats=F)



ggplot(simresults_compiled%>%filter(incidence%in%c(6,16,26,36),cohort_size%in%c(100,300,500,900)),aes(x=factor(or1),y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_uq/10))+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Odds ratio of GOI and PC1")+facet_grid(cohort_size~incidence)+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(plot.title = element_text(hjust = 0.5))

ggplot(simresults_compiled%>%filter(incidence%in%c(6,16,26,36),cohort_size%in%500),aes(x=factor(or1),y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_uq/10))+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Odds ratio of GOI and PC1")+facet_wrap(~incidence,ncol=4)+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(plot.title = element_text(hjust = 0.5),axis.text.x=element_text(angle=90,hjust=.5,vjust=.5),axis.text=element_text(face="bold",size="9",color="black"))
# ggsave("score_heatmap_bestoption1.pdf",width=8,heigh=2.5,units="in",useDingbats=F)
# sort(unique(simresults_compiled$or2))

ggplot(simresults_compiled%>%filter(cohort_size%in%500),aes(x=factor(or1),y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_median/10))+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Odds ratio of GOI and PC1")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(plot.title = element_text(hjust = 0.5))
# ggsave("score_heatmap_bestoption.pdf",width=4,heigh=3,units="in",useDingbats=F)

ggplot(simresults_compiled%>%filter(or1%in%.3,or2%in%.09),aes(x=incidence/cohort_size,y=goipc1_isgreater_median/10))+geom_point()
ggplot(simresults_compiled%>%filter(or1%in%.1,or2%in%.05,cohort_size%in%500),aes(x=incidence,y=goipc1_isgreater_median/10))+geom_point()
  # scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Odds ratio of GOI and PC1")+
  # scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(plot.title = element_text(hjust = 0.5))
# or1%in%c(.05,0.1,.5),cohort_size%in%c(100,500,900)
ggplot(simresults_compiled%>%filter(cohort_size%in%c(100,300,500)),aes(x=incidence,y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_median/10))+facet_grid(cohort_size~or1)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_continuous(expand = c(0,0),name="Incidence")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1vsPC2")+ggtitle("Odds ratio of GOIvsPC1")+
  theme(plot.title = element_text(hjust = 0.5))

```
  
  
### Here, I am looking at whether frequency correction of the positive control genes is in fact a more fair test of a comparison with gene of interest than no frequency correction. i.e. does it matter if the positive control genes are downsampled to the frequency of the gene of interest? Answer: yes. If the positive control genes were present at a lower frequency, it would be tougher to tell PC1 vs PC2 apart from GOI vs PC1.
```{r}
ggplot(simresults_compiled%>%filter(or1<=.7,or1>=.2),aes(x=incidence,y=or2))+geom_tile(color="black",aes(fill=fp_corrected_95))+facet_grid(cohort_size~or1)+scale_fill_gradient2(low ="red" ,mid ="white",high ="blue",name="Regions where downsampling helps")

ggplot(simresults_compiled%>%filter(or1<=.7,or1>=.2),aes(x=incidence,y=or2))+geom_tile(color="black",aes(fill=fp_corrected_99))+facet_grid(cohort_size~or1)+scale_fill_gradient2(low ="red" ,mid ="white",high ="blue",name="Regions where downsampling helps")
```




Choosing plots for putting in paper
```{r}
ggplot(simresults_compiled%>%filter(cohort_size%in%500),aes(x=factor(or1),y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_uq/10))+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Odds ratio of GOI and PC1")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(plot.title = element_text(hjust = 0.5))+theme(plot.title = element_text(hjust = 0.5),axis.text.x=element_text(angle=90,hjust=.5,vjust=.5),axis.text=element_text(face="bold",size="9",color="black"))
# ggsave("score_heatmap_bestoption.pdf",width=4,heigh=3,units="in",useDingbats=F)

ggplot(simresults_compiled%>%filter(or1%in%c(.05,.3,.6),incidence%in%c(4,8,12,16,20,24,28,32,36),cohort_size%in%500),aes(x=factor(incidence),y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_median/10))+facet_wrap(~or1,ncol=6)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Incidence")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(strip.text=element_blank(),plot.title = element_text(hjust = 0.5),axis.text.x=element_text(angle=90,hjust=.5,vjust=.5),axis.text=element_text(face="bold",size="9",color="black"),legend.position = "none")

# ggsave("score_heatmap_bestoption1.pdf",width=6,heigh=2.5,units="in",useDingbats=F)

ggplot(simresults_compiled%>%filter(or1%in%c(.1,.2,.5),incidence%in%c(4,8,12,16,20,24,28,32,36),cohort_size%in%500),aes(x=factor(incidence),y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_raw_median/10))+facet_wrap(~or1,ncol=6)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Incidence")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(strip.text=element_blank(),plot.title = element_text(hjust = 0.5),axis.text.x=element_text(angle=90,hjust=.5,vjust=.5),axis.text=element_text(face="bold",size="9",color="black"),legend.position = "none")
# ggsave("score_heatmap_bestoption1.pdf",width=6,heigh=2.5,units="in",useDingbats=F)

ggplot(simresults_compiled%>%filter(cohort_size%in%500,incidence%in%c(4,8,12,16,20,24,28,32,36)),aes(x=factor(incidence),y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_raw_median/10))+facet_wrap(~or1,ncol=4)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Incidence")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")+theme(plot.title = element_text(hjust = 0.5),axis.text.x=element_text(angle=90,hjust=.5,vjust=.5),axis.text=element_text(face="bold",size="9",color="black"),legend.position = "none")
# ggsave("score_heatmap_bestoption1_supplement.pdf",width=6,heigh=6,units="in",useDingbats=F)

ggplot(simresults_compiled%>%filter(or1%in%c(.05,.1,.5),cohort_size%in%c(100,300,500)),aes(x=incidence,y=or2))+geom_tile(color="black",aes(fill=goipc1_isgreater_median/10))+facet_grid(cohort_size~or1)+scale_fill_gradient2(low ="red" ,mid ="white",midpoint=50,high ="blue",name="Score")+scale_x_discrete(expand = c(0,0),name="Incidence")+
  scale_y_continuous(expand = c(0,0),name="Odds ratio of PC1 and PC2")


# a=simresults_compiled%>%filter(cohort_size%in%500,or1%in%.7,or2%in%.05,incidence%in%)
```



```{r,eval=F}
# simresults_subset=simresults_compiled%>%filter(or2%in%0.01,or1%in%c(0.01,.1,1))
simresults_subset=unlist(simresults_compiled_alldata)

simresults_unlisted=data.frame(unlist(lapply(simresults_compiled_alldata,'[[',1)),
             unlist(lapply(simresults_compiled_alldata,'[[',2)),
             unlist(lapply(simresults_compiled_alldata,'[[',3)),
             unlist(lapply(simresults_compiled_alldata,'[[',4)))
simresults_unlisted$list1=(lapply(simresults_compiled_alldata,'[[',5))
simresults_unlisted$list2=(lapply(simresults_compiled_alldata,'[[',6))

colnames(simresults_unlisted)=c("cohort_size","incidence","or1","or2","or1_list","or2_list")

simresults_unlisted=simresults_unlisted%>%filter(or2%in%0.01,or1%in%c(0.05,.1,1),cohort_size%in%500,incidence%in%c(4,8,12,16,20))
# library(reshape2)
a=simresults_unlisted%>%filter(incidence%in%8,or2%in%0.01,or1%in%.05)
# b=unnest(a)
median(b$or2_list)>median(b$or1_list)

# library(tidyr)
simresults_melted=unnest(simresults_unlisted)
simresults_melted2=melt(simresults_melted,
                        id.vars = c("cohort_size","or1","or2","incidence"),
                        measure.vars =c("or1_list","or2_list"),
                        variable.name = "Comparison",
                        value.name = "OR"
                        )
ggplot(simresults_melted2,aes(x=factor(incidence),y=OR,fill=Comparison))+facet_wrap(~or1)+geom_boxplot()+scale_y_continuous(trans="log2")+cleanup

plotly=ggplot(simresults_melted2,aes(x=factor(incidence),y=OR,fill=Comparison))+facet_wrap(~or1)+geom_boxplot()+cleanup
ggplotly(plotly)




simresults_unlisted=data.frame(unlist(lapply(simresults_compiled_alldata,'[[',1)),
             unlist(lapply(simresults_compiled_alldata,'[[',2)),
             unlist(lapply(simresults_compiled_alldata,'[[',3)),
             unlist(lapply(simresults_compiled_alldata,'[[',4)))
simresults_unlisted$list1=(lapply(simresults_compiled_alldata,'[[',5))
simresults_unlisted$list2=(lapply(simresults_compiled_alldata,'[[',6))

colnames(simresults_unlisted)=c("cohort_size","incidence","or1","or2","or1_list","or2_list")
simresults_unlisted=simresults_unlisted%>%filter(or1%in%.5,or2%in%c(0.05,.11,.19),cohort_size%in%500,incidence%in%c(4,8,12,16,20))

simresults_melted=unnest(simresults_unlisted)
simresults_melted2=melt(simresults_melted,
                        id.vars = c("cohort_size","or1","or2","incidence"),
                        measure.vars =c("or1_list","or2_list"),
                        variable.name = "Comparison",
                        value.name = "OR"
                        )

ggplot(simresults_melted2,aes(x=factor(incidence),y=OR,fill=Comparison))+facet_wrap(~or2)+geom_boxplot()+scale_y_continuous(trans="log2")+cleanup

plotly=ggplot(simresults_melted2,aes(x=factor(incidence),y=OR,fill=Comparison))+facet_wrap(~or2)+geom_boxplot()+cleanup
ggplotly(plotly)
# sort(unique(simresults_unlisted$or2))

#I used the following website to make sure that my Gaussian elimination was working properly
# https://www.emathhelp.net/calculators/linear-algebra/gauss-jordan-elimination-calculator/?i=%5B%5B1%2C1%2C0%2C0%2C.4836%5D%2C%5B1%2C1%2C1%2C1%2C1%5D%2C%5B0%2C1%2C0%2C-1%2C0%5D%2C%5B1%2C0%2C-0.01%2C0%2C0%5D%5D&reduced=on

```

